### 2013-11-20 Paper.tex
First commit

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[font=scriptsize,labelfont={bf}]{caption}
\usepackage[justification=raggedright,nearskip=10pt,farskip=0pt]{subfig}
\usepackage{calc}
\usepackage{url}
\usepackage{array}
\usepackage{rotating}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{layouts}
\usepackage{titletoc}
\usepackage{setspace}
\usepackage{longtable}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{footnote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\includegraphic[#1]#2{}
\def\includegraphics[#1]#2{}

%%% Put your definitions there:
\startlocaldefs
\hbadness=9999999 %avoid complaining for underfull h/vbox
\vbadness=9999999

% \setlength{\columnwidth}{86mm}
 \captionsetup[subfloat]{position=top, justification=raggedright, singlelinecheck=false, labelformat=simple, font={scriptsize,sf,bf}}
 \renewcommand*\thesubfigure{\Alph{subfigure}}
 \setlength{\belowcaptionskip}{-3mm}
 %\setlength{\abovecaptionskip}{0mm}
%\addtolength{\abovecaptionskip}{-3mm}

%\subfiguretopcaptrue

\pdfminorversion=4
\pdfobjcompresslevel=0

\pgfdeclareimage[height=3ex]{win-logo}{images/win_logo}
\pgfdeclareimage[height=3ex]{mac-logo}{images/mac_logo}
\pgfdeclareimage[height=3ex]{linux-logo}{images/linux_logo}
\pgfdeclareimage[height=3ex]{ubuntu-logo}{images/ubuntu_logo}
\pgfdeclareimage[height=2ex]{mac-command}{images/mac_command}
\pgfdeclareimage[height=2ex]{lower-bound}{images/lowerbound}
\pgfdeclareimage[height=2ex]{upper-bound}{images/upperbound_green}

\newcommand{\winsymbol}{\raisebox{-0.8ex}{\pgfuseimage{win-logo}}}
\newcommand{\macsymbol}{\raisebox{-0.5ex}{\pgfuseimage{mac-logo}}}
\newcommand{\linuxsymbol}{\raisebox{-0.8ex}{\pgfuseimage{linux-logo}}}
\newcommand{\ubuntusymbol}{\raisebox{-0.9ex}{\pgfuseimage{ubuntu-logo}}}
\newcommand{\winmaclinux}[3]{\begin{itemize}
    \item[\winsymbol] #1
    \item[\macsymbol] #2
    \item[\linuxsymbol] #3
\end{itemize}}
\newcommand{\macpc}[2]{\begin{itemize}
    \item[\macsymbol] #1
    \item[\winsymbol\ \  \linuxsymbol] #2
\end{itemize}}

\newcommand{\maccmd}{\raisebox{-0.35ex}{\pgfuseimage{mac-command}}}

\setlength{\abovecaptionskip}{1.3mm}
%\setlength{\belowcaptionskip}{1mm}

\def\ta{Timed Automaton}
\def\tas{Timed Automata}

%\setvaluelist{bmcsymbol}{*,\textdagger,\^{}}
\def\textdagger{\#{}} %I would be very happy to avoid the distasteful symbol \textdagger (or \dagger, whatever).

\endlocaldefs

\makeindex

\begin{document}

\begin{frontmatter}

\begin{fmbox}
\dochead{Software}
% \addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}
% \firstpage{1}

\title[Bringing biological networks to life with ANIMO]{Bringing biological networks to life with ANIMO}

% \author[Schivo, Scholma \textit{et~al}]{Stefano Schivo\,$^{1,*}$, Jetse Scholma\,$^{2,*}$, Brend Wanders\,$^3$, Ricardo A. Urquidi Camacho\,$^2$, Paul E. van der Vet\,$^3$, Marcel Karperien\,$^2$, Rom Langerak\,$^1$, Jaco van de Pol\,$^{1,\ddag}$ and Janine N. Post\,$^{2,\ddag}$}
% \address{$^{1}$Formal Methods and Tools, Faculty of EEMCS, University of Twente, 7522NH Enschede, The~Netherlands;
% $^{2}$Developmental BioEngineering, MIRA Institute for Biomedical Technology and Technical Medicine, University of Twente, 7522NH Enschede, The~Netherlands;
% $^{3}$Human Media Interaction, Faculty of EEMCS, University of Twente, 7522NH Enschede, The~Netherlands}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\dagger{AA}
\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   %corref={aff1},                       % id of corresponding address, if any
   noteref={n1},                        % id's of article notes, if any
   email={s.schivo@utwente.nl}   % email address
]{\inits{S}\fnm{Stefano} \snm{Schivo}}
\author[
   addressref={aff2},
   noteref={n1},
   email={j.scholma@utwente.nl}
]{\inits{J}\fnm{Jetse} \snm{Scholma}}
\author[
   addressref={aff3},
   email={b.wanders@utwente.nl}
]{\inits{B}\fnm{Brend} \snm{Wanders}}
\author[
   addressref={aff2},
   email={rauc6788@hotmail.com}
]{\inits{RA}\fnm{Ricardo A.} \snm{Urquidi Camacho}}
\author[
   addressref={aff3},
   email={p.e.vandervet@utwente.nl}
]{\inits{PE}\fnm{Paul E.} \snm{van der Vet}}
\author[
   addressref={aff2},
   email={h.b.j.karperien@utwente.nl}
]{\inits{M}\fnm{Marcel} \snm{Karperien}}
\author[
   addressref={aff1},
   email={r.langerak@utwente.nl}
]{\inits{R}\fnm{Rom} \snm{Langerak}}
\author[
   addressref={aff1},
   %corref={aff1},
   email={j.c.vandepol@utwente.nl}
]{\inits{J}\fnm{Jaco} \snm{van de Pol}}
\author[
   addressref={aff2},
   corref={aff2},
   email={j.n.post@utwente.nl}
]{\inits{JN}\fnm{Janine N.} \snm{Post}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgname{Formal Methods and Tools, Faculty of EEMCS, University of Twente}, % university, etc
  %\street{Hallenweg},                     %
  \postcode{7522NH}                           % post or zip code
  \city{Enschede},                              % city
  \cny{The~Netherlands}                                    % country
}
\address[id=aff2]{%
  \orgname{Developmental BioEngineering, MIRA Institute for Biomedical Technology and Technical Medicine, University of Twente},
  %\street{ },
  \postcode{7522NH}
  \city{Enschede},
  \cny{The~Netherlands}
}
\address[id=aff3]{%
  \orgname{Human Media Interaction, Faculty of EEMCS, University of Twente},
  %\street{ },
  \postcode{7522NH}
  \city{Enschede},
  \cny{The~Netherlands}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}
% 
% \let\oldthefootnote\thefootnote
% \renewcommand{\thefootnote}{\fnsymbol{footnote}}
% \footnotetext[1]{These authors contributed equally to this work.}
% \footnotetext[3]{To whom correspondence should be addressed.}
% \let\thefootnote\oldthefootnote

\end{fmbox}% comment this for two column layout

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract <= 350 words. Headers: background, results, conclusions
\parttitle{Motivation}
Computational modelling is a vital part of the empirical cycle,
formalizing existing knowledge and experimental data in a systematic way.
Existing methods are often parameter intensive, or supporting tools are
not accessible enough to the biological community. These factors
hamper widespread application of modelling in biological research.
\parttitle{Results}
We present an approach to modelling based on a series of
abstractions that limit the number of parameters in the model.
To support this approach, we developed a user-friendly software tool,
ANIMO (Analysis of Networks with Interactive MOdelling).
An ANIMO model of signal transduction events downstream of TNF$\alpha$
and EGF in HT-29 human colon carcinoma cells is described as a case study.
This model gives a formal description of crosstalk between the pathways at
different cellular levels and led to the formulation of novel testable hypotheses.
\parttitle{Availability and implementation}
ANIMO is implemented as a plug-in to the network visualization tool Cytoscape
and can be downloaded from \url{http://fmt.cs.utwente.nl/tools/animo}.
Additional information and a step-by-step installation guide can be found
in Supplementary Section~\ref{sec:animo-installation}. A user manual is 
provided in Supplementary Section~\ref{sec:animo-manual}.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{sample}
\kwd{article}
\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

\section*{Introduction}\label{sec:introduction}
In living cells, processes are regulated by networks of interacting molecules.
Aberrations in these networks underlie a wide range of pathologies. The development of new therapies
requires a thorough insight in the functioning of these networks. Obtaining such insight can be a challenging task.
Feedback loops and crosstalk between pathways lead to an intricate wiring of the network.
Hence, it is necessary to study the ensemble of molecules involved, because the 
behaviour of individual molecules is not sufficient for a complete understanding. 
Since the human brain is ill-suited to grasp the non-linear dynamics of these complex networks and
the entailed emergent properties, the role of computational support is increasing
in molecular biology.

The systems biology approach to understanding biological systems starts off from a
scientific question and then follows an empirical cycle \--\ or rather a positive spiral \--\ of
knowledge/theory $\rightarrow$ model $\rightarrow$ hypotheses $\rightarrow$ experiments $\rightarrow$
observations $\rightarrow$ update and/or refinement of knowledge/theory,
until an answer to the original question is found (Figure~\ref{fig:empirical-spiral}).
A model plays a pivotal role in this cycle:
\begin{enumerate}
  \item to organize data and store knowledge,
  \item to structure reasoning and discussion
  \item to perform \emph{in silico} experiments and derive hypotheses.
\end{enumerate}
An \emph{in silico} model is always a simplified representation of biological reality and is never the 
aim in itself.
Rather, it is a powerful means in the process of gaining an understanding of the biological system.
Given its role in the empirical cycle, the process of modelling is especially effective
when applied by the experts with respect to a certain biological system. Biologists usually have a good sense 
of cause-and-effect relationships of molecular interactions. In addition, they are the most knowledgeable
on the network topology and the dynamics of the biological system they are studying.
Since they also benefit most from the generation of hypotheses and from an efficient experimental design, 
biologists would be the primary candidates to construct models of their research topic.

As models are a formalization of knowledge or theories, an underlying formalism is needed to express
this knowledge. Different formal methods have been successfully applied to construct representations
of biological systems. Among these methods are Boolean logic~\cite{boolean-networks-flower,boolean-networks2},
ordinary differential equations (ODEs, reviewed by~\cite[]{hidde-review}),
interacting state machines~\cite{interacting-sm1,interacting-sm2},
process calculi~\cite{blenx,bio-pepa}, Timed Automata~\cite{ta-siebert,bartocci-oscillators,
oded-ode-ta-discretization} and Petri nets~\cite{petri-nets,petri-nets2}.
Most of these formal methods have been implemented into software tools to aid the process
of modelling. Due to the lack of such a supporting tool, \tas\ have remained a less 
frequently applied method.

\tas\ have been developed to model the dynamic behaviour of systems with processes running in parallel~\cite{timed-automata-alur-dill}. 
As such, \tas\ have been applied in communication protocols and industrial control engineering~\cite{ta-audio-protocol,ta-wap-gateway,ta-ws-bap}. The 
parallels between these application areas and regulatory processes in cells have triggered the step towards 
their use in biology.
\cite{ta-siebert} use \tas\ to extend a classical modelling paradigm
~\cite{thomas-formalism}, allowing to add temporal dynamics to gene network models.
\cite{bartocci-oscillators} describe a model of biological oscillators and test 
synchronization properties in this dynamic system.
A discretization of ODEs to \tas\ is proposed by~\cite{oded-ode-ta-discretization}, applying
a translation between the two formalisms to an example gene regulatory network. Two 
different approaches to transforming
a Petri net model into \tas\ are presented by~\cite{ta-giapponesi},
who also address the important issue of state space explosion in their paper.
Finally, \cite{ta-radiazioni} propose an \emph{ad hoc} \tas\ model of a radiation treatment
system, which is then validated through UPPAAL.\\
Each of these approaches has been successfully validated, demonstrating the potential of \tas\
in biological applications. However, these approaches were all limited to simple
or specific examples and none of these modelling methods
has led to a tool implementation of the proposed method to encourage a broader use
of \tas\ in molecular biology.

Mastery of most existing modelling tools requires training and experience in mathematical modelling. 
In this respect, a lack of tradition in quantitative
reasoning and formal methods within the biological community at large is still a stumbling block for
widespread application of modelling of biological systems. Here, we present an intuitive method for the
construction of formal in silico models of the dynamics of molecular networks, supported by a novel,
user friendly modelling tool, ANIMO (Analysis of Networks with Interactive 
MOdelling,~\cite[]{animo-bibe}). \tas\ are used as the underlying mathematical formalism.

In the Methods section, we will explain how choosing a suitable abstraction level can make the construction
of models more intuitive. We will then show how ANIMO is designed to support the modelling process following
this approach. Construction of a small model based on experimental data will exemplify the method that we
propose. In the Results section, we first show an ANIMO model of the genes and proteins that constitute
the circadian clock network in \emph{Drosophila Melanogaster}. The remainder of that section is dedicated to
illustrate how a single modelling iteration in the empirical cycle is used to compile prior knowledge and
experimental data into a model, perform \emph{in silico} simulations and derive meaningful testable 
hypotheses. These hypotheses are supported by literature on interactions in different cell types.

\section*{Methods}
\subsection*{Modelling abstractions}\label{subsec:abstractions}

In living cells, cascades of chemical and physical interactions enable propagation of signals through molecular networks.
In this process, the activity of upstream molecules induces a change in the 
concentration or activity of downstream molecules. For many reactions, the values of the kinetic parameters 
are unknown or difficult to collect. This lack of knowledge hampers the feasibility 
of computational models that describe molecular networks in fine mechanistic detail, especially for larger networks.
As a solution to this problem, we propose the construction of models at a higher level of abstraction, 
thereby reducing the number of parameters involved. In choosing a suitable abstraction level, it is important to 
retain enough descriptive power to give a meaningful formal description of the topology and the 
associated dynamic behaviour of biological networks.

As a first abstraction in ANIMO models, the active and inactive forms of each network component 
are represented together by a single node in the network.
Each of these nodes is characterized by its \emph{activity level}, which
represents the fraction of active molecules of that molecular species. When a molecule is known to be 
constitutively active, changes in concentrations of that molecule are treated as changes in its activity level.
Activity levels are discretized into integer variables with a user-defined granularity, ranging from 
Boolean (2 levels) to near-continuous (100 levels).

Detailed biochemical reaction mechanisms are abstracted to \emph{interactions}, which can 
represent either activations ($\rightarrow$) or inhibitions ($\dashv$\hspace{0.1em}). 
This aggregation of elementary reactions into single interaction steps reduces the number of kinetic 
parameters involved, while preserving cause-and-effect relationships.
For example, consider a reaction in which enzyme $E$ phosphorylates and activates substrate $S$, 
transferring a phosphate group from a molecule of ATP to a molecule of $S$. Biochemically, this reaction 
can be represented as
$$
\mbox{\it E} + \mbox{\it S} + \mbox{\it ATP} \rightleftarrows \mbox{\it ES} + \mbox{\it ATP} \rightarrow \mbox{\it ES}^{\mbox{\scriptsize \it P}} + \mbox{\it ADP} \rightleftarrows \mbox{\it E} + \mbox{\it S}^{\mbox{\scriptsize\it P}} + \mbox{\it ADP},
$$
with conservation condition $\mbox{\it S} + \mbox{\it S}^{\mbox{\scriptsize\it P}} = \mbox{constant}$ and $\mbox{\it ATP} + \mbox{\it ADP} = \mbox{constant}$.\\
Under the assumption of ATP constantly being replenished by the cell, this reaction is abstracted in ANIMO to the corresponding interaction
$$
\mbox{\it E} \rightarrow \mbox{\it S}.
$$
Each occurrence of the interaction $E \rightarrow S$ will increase the activity level of $S$ by one discrete step. 
Since the activity level is defined as the active fraction of a molecular species, an increase in the active fraction
implies a decrease in the inactive fraction. Hence, the original conservation condition is automatically  
satisfied.
The interaction rate, $R$, depends on the activity levels of the reactants involved and on a single kinetic
parameter $k$ that is set by the user. 
The three available interaction scenarios can be interpreted as abstracted kinetic rate laws:
\begin{enumerate}
  \item $R = k \times [E]$: the interaction rate depends only on the activity level of the upstream node.
  \item $R = k \times [E] \times [1 - S]$ (activations) or $R = k \times [E] \times [S]$ (inhibitions): the rate 
  depends on the activity levels of both the upstream and downstream participants. Activations depend on the 
  presence of inactive substrate, \emph{[1 - S]}, whereas inhibitions depend on the level of active substrate,
  \emph{[S]}.
  \item $R = k \times [E_1] \times [E_2]$: this scenario can be used when the activation or inhibition
  of a downstream node depends on the simultaneous activity of two upstream nodes. This scenario is comparable to an
  \emph{AND-gate} in Boolean logic.
\end{enumerate}
We will show in Section~\ref{sec:results} that the abstraction proposed here preserves ample
expressivity to capture the dynamic behaviour of a biological network.

\subsection*{Modelling interactions with Timed Automata}\label{subsec:timed-automata}
\def\ta{TA}
\def\tas{TA}

Timed Automata have been shown to be a powerful formalism to model biological processes
~\cite{ta-siebert,bartocci-oscillators,oded-ode-ta-discretization}. A timed automaton consists of locations
and transitions between these locations (see Fig.~\ref{fig:abstraction-mek-erk}), and a system of timed automata can be 
used to model a system of interacting molecules. At any time, each automaton is in a specific location, and together 
these locations represent the current state of the biological system. Each timed automaton can have one or more local clocks
associated to it, allowing temporal control of transitions between locations. These transitions are used to 
represent interactions between molecules. Fast interactions take less time than slow interactions 
to perform an activation or inhibition step. We have previously described in detail how the 
scenarios presented in Section~\ref{subsec:abstractions} can be used to calculate the timing of molecular 
interactions to give a description of network dynamics~\cite{animo-bibe}. Figure~\ref{fig:abstraction-mek-erk}
presents a small example that illustrates the basic properties of \tas. 
This model describes the activation of ERK by MEK\footnote{All acronyms used in this paper
and their corresponding UniProt IDs are listed in Suppl. Sect.~\ref{suppl-sec:names}.}.

\subsection*{ANIMO}
The modelling approach described in Section~\ref{subsec:abstractions} and Section~\ref{subsec:timed-automata}
is implemented in the
software tool ANIMO (Analysis of Networks with Interactive MOdelling, \cite{animo-bibe}) as
a plug-in to the network visualization tool Cytoscape~\cite{cytoscape}. The visual interface of Cytoscape
makes the construction, expansion and rewiring of a network topology a fast and user-friendly process.

When a new node is added to the network, it has to be initialized with the number of activity levels and 
its initial activity. 
For each interaction, a scenario needs to be selected, together with the corresponding kinetic parameter and 
the interaction type: activation or inhibition. All settings can be readily adapted by double clicking, or via a 
table of nodes or interactions.

ANIMO automatically translates the user input to a \tas\ model, which is then simulated with the model 
checking tool UPPAAL~\cite{uppaal}. The results are subsequently parsed and translated to a graph that shows
the dynamic behaviour of nodes in the network. 
A schematic overview of this process is given in Supplementary Section~\ref{suppl-sec:animo-ta}.
No training or prior knowledge on the use of \tas\ or UPPAAL is needed in order to benefit from ANIMO.
Nevertheless, the \tas\ model and the model checking process in UPPAAL can be accessed when desired by the user.

The dynamic behaviour of a model can be interactively explored by
moving a time slider underneath the graph to highlight time points in a simulation. In the network view,
each node will be coloured according to its activity level at the selected time point. 
Experimental data can be compared to the model by importing and superposing these data 
upon an output graph from the model (Figure~\ref{fig:small-model} B,D,F). The ANIMO user workflow and the 
features described above are illustrated in Suppl. Video 1.

\subsection*{Using ANIMO to build a model based on data}\label{subsec:case-study}
To illustrate the process of modelling with ANIMO, we show the construction of a small model based on a 
literature compendium of signal transduction events in HT-29 human colon carcinoma cells~\cite{pathway-compendium}. 
This data set comprises triplicate
measurements of 11 different protein activities or post-translational modification states at 13 time points after
treatment with different combinations of TNF$\alpha$, EGF and insulin.
The data set contains relative protein levels and activities, which is a typical situation in biochemistry.

As a first step, we normalized measurements for each protein to the
maximum value in the complete experiment. This normalization results in a nondimensionalized data set that 
is suitable for use with ANIMO.

In Figure~\ref{fig:small-model}, we show the stepwise construction of a model of a small part of the network that is
able to account for measured variations in activity of IKK, JNK1, MK2, Casp8 and Casp3 upon stimulation with 100~ng/ml TNF$\alpha$.
In this example we aimed for inclusion of a minimum number of nodes in the network, 
while preserving biological relationships.
Multi-step cascades were aggregated into a single step when possible. Parameters for all interactions were 
set manually, and a close match was obtained between the model and the patterns that are present in the dataset.

A more comprehensive model based on the same dataset is presented in Section~\ref{subsec:case-study-larger}.

\section*{Results}\label{sec:results}
\subsection*{Expressivity of ANIMO models}\label{sec:animo-drosophila}
Results obtained with ANIMO are comparable to results with other modeling
approaches. To demonstrate this, Figure~\ref{fig:drosophila-model}
represents an ANIMO model of the circadian clock in \emph{Drosophila Melanogaster}, based on the work
by~\cite{drosophila-ode-model}, where ordinary differential equations (ODEs) were used.
The cyclic behavior of the circadian clock is based on the alternating formation and destruction of the
CYC/CLK protein complex.
Concentration levels of this complex are in turn regulated by a series of proteins which are produced as
a consequence of CYC/CLK formation. The CWO protein
is central to the functioning of the network, as it degrades the mRNA for most of the involved proteins.
As such, CWO act as an inhibitor that counterbalances the effect of CYC/CLK.
The positive influence of the light-regulated cryptochrome CRY on the degradation of TIM is a consequence
of the passage between day and night, allowing
the circadian clock to synchronize to a time zone (see Suppl. Sect.~\ref{suppl:repressilator}).

The output of the ANIMO model in Figure~\ref{fig:drosophila-model} closely matches the original ODE model.
In particular, the oscillations in both models show the same periods and phases (see Suppl. Fig~\ref{suppl-fig:grafici-drosophila}).
Due to the compositional nature of \tas\, ANIMO allows for intuitive \emph{in silico} knock-out experiments,
by right-clicking a node in the model and disabling it. Such experiments have been done
before~\cite{drosophila-ode-model} and give similar results in our model.
% Details of the comparison
% between the ANIMO model and the original ODE model is given in Supplementary Section~\ref{suppl-sec:animo-drosophila}.

\subsection*{Case study: using ANIMO to generate hypotheses}\label{subsec:case-study-larger}
In order to validate our modeling approach,
we constructed a larger model of the signaling network downstream of TNF$\alpha$ and EGF, formalizing
the crosstalk that takes place between the pathways at different levels of cellular regulation.
We first modeled the two pathways in isolation~(Suppl. Figs.~\ref{fig:large-model-tnf}, \ref{fig:large-model-egf}),
using information on protein interactions from
the KEGG~\cite{kegg} and phosphosite~\cite{phosphosite} databases. These models were fitted to experimental data
from studies by~\cite{pathway-compendium} and~\cite{pathway-autocrine}.
We then merged the two pathways into a single model and added autocrine crosstalk between the pathways that has been
described by~\cite{pathway-autocrine}.
Briefly, stimulation with TNF$\alpha$ leads to a rapid release of TGF$\alpha$ ({\sf TGFa} in the model),
which activates the EGF receptor ({\sf EGFR}).
This activation causes secretion of IL-1$\alpha$ ({\sf IL-1a}) at later time points.
The effect of IL-1$\alpha$ is down-regulated by the secretion of IL-1 receptor antagonist ({\sf IL-1ra})
downstream of TNF$\alpha$.
The resulting model (Fig.~\ref{fig:large-model-no-hypotheses}) was compared to the experimental data
for treatments with 100 ng/ml TNF alone and 100 ng/ml EGF alone (data not shown)~\cite{pathway-compendium}.

At this point, the behavior of the model deviated from the data for some of the nodes.
This is an interesting situation, as it requires
modifications to the model, that can be interpreted as new hypotheses. Below, we give two examples and show how
adaptation of the model can be used to generate novel testable hypotheses.

Experimentally,
treatment with TGF$\alpha$ alone does not lead to secretion of IL-1$\alpha$. Instead, a co-stimulation with
TGF$\alpha$ and TNF$\alpha$ is required~\cite{pathway-autocrine}.
However, in the first version of the model, treatment with TGF$\alpha$ was sufficient for
IL-1$\alpha$ expression (Fig.~\ref{fig:large-model-graph1}). Given the time delay until secretion of IL-1$\alpha$, it can be
expected that \emph{de novo} synthesis of IL-1$\alpha$ is required and that both
TNF$\alpha$ and TGF$\alpha$ are needed to activate transcription of the IL-1$\alpha$ gene.
JNK1 and ERK signal downstream of TNF$\alpha$ and TGF$\alpha$, respectively, and are known
to affect the activity of multiple transcription factors. We altered the model to make
activation of IL-1$\alpha$ expression dependent on both JNK1 activity and ERK activity
(Suppl. Fig.~\ref{fig:large-model-hypotheses}, arrows linking {\sf JNK1} and {\sf ERK} to {\sf IL-1a gene}).
After this modification to the model, IL-1$\alpha$ was no longer secreted
upon stimulation with TGF$\alpha$ alone, which greatly improved the fit between the measured IL-1$\alpha$
levels and the model (Fig.~\ref{fig:large-model-graph2}). This hypothesis could now be used to
design a new experiment to validate IL-1$\alpha$ as a target of combined JNK1 activity and ERK activity in
HT-29 cells. For example, kinase inhibitors specific to JNK1 and ERK could be used to confirm that activity of
both kinases is required for expression and secretion of IL-1$\alpha$. Performing the experiment is beyond
the scope of this study, but this hypothesis finds support in literature.
Transcription factors c-Jun and c-Fos together
form a heterodimer known as AP-1 and are activated by JNK1 and ERK,
respectively~\cite{jnk-signaling,cfos-cjun}. AP-1 has been reported to bind to the
promoter of IL-1$\alpha$, providing evidence for a role in the regulation of IL-1$\alpha$
expression~\cite{ap1-il1a}. Based on these findings in literature we included c-Jun and
c-Fos in our model as transcriptional activators of IL-1$\alpha$ (Fig.\ref{fig:large-model-complete}).

As a second example, we considered the behaviour of JNK1 and MK2. In the model, both
proteins were located downstream of TNF$\alpha$ but not TGF$\alpha$ or EGF. Hence, the
model did not show an effect of C225, a pharmacological inhibitor of ligand-EGFR
binding, on activation of JNK1 or MK2 after stimulation with TNF$\alpha$. However, experimental
data show that C225 strongly reduces activation of JNK1 and MK2 upon stimulation with
TNF$\alpha$~\cite{pathway-autocrine}.
This fact is indicative of a role for EGFR in activation of JNK1 and MK2. Since both JNK1 and MK2
are located downstream of MEKK1, we hypothesized that activation
of MEKK1 is dependent on
both TNF$\alpha$-signalling and TGF$\alpha$-signalling. In the model we added a new
hypothetical node {\sf Hyp~2} (hypothesis~2) to link EGFR to MEKK1 (Suppl. Fig.~\ref{fig:large-model-hypotheses}).
This addition led to an improved fit of the model to the data upon treatment with TNF$\alpha$ + C225:
activation of both MK2 and JNK1 was strongly suppressed by C225 (Fig.~\ref{fig:large-model-graph4}).
Stimulation with EGF alone did not lead to activation of JNK1 and MK2.
These data support the validity of the modification to the model.
Further support for a link between EGFR and MEKK1 was found in literature. Specifically,
Ras has been reported as a direct activator of
MEKK1~\cite{ras-mekk1}. EGFR is a well-known and potent activator of Ras,
which is why it was already in our network~\cite{kegg}.
Other studies also report activation of JNK1 and phosphorylation of c-Jun downstream of Ras, which is consistent with
an interaction between Ras and MEKK1~\cite{cfos-cjun,ras-jnk1}.
Based on these findings, we adapted
our model by removing the {\sf Hyp~2} node and creating a direct interaction between Ras
and MEKK1 (Fig.~\ref{fig:large-model-complete}). Experimentally, the role of Ras could be confirmed by using a
pharmacological inhibitor of Ras activity, and measuring the effect of this inhibitor on the activation of JNK1 and MK2.
Together, our model suggests that EGFR activity is required
but not sufficient for activation of JNK1 and MK2 in HT-29 cells.

There are other nodes for which the experimental data deviates from the model in one or more of the experimental conditions.
A comparison between model and experimental data can be found in Figures~\ref{fig:differences1}, \ref{fig:differences2} and~\ref{fig:differences3}.
A complete deciphering of the signalling events
in this biological system is outside the scope of this paper. Instead, we illustrated how interactive modelling of the dynamic behaviour
of a signal transduction network can be used to extend previous pathway topologies and can lead to the generation of novel hypotheses.

\section*{Discussion}
The ultimate aim of research projects is to solve a problem or get
the answer to a scientific question. In biology, an in-depth understanding of
the relevant biological system is an important step towards this goal. Successive 
repetitions of the empirical cycle result in a stepwise increase in understanding,
until the goal is reached. For complex biological systems, computational modelling is 
indispensable in this process. The mere act of creating 
a computational model based on prior knowledge, experimental data and hypotheses 
assists in gaining more insight in the system.

%Furthermore, \emph{in silico} experiments 
%with the model can be used to make predictions that cannot be made by the human brain.

We developed a new modelling approach by proposing a series of abstractions from the detailed 
molecular mechanisms of biological systems. These abstractions reduce the need for kinetic 
parameters, while preserving enough expressivity for a useful description of the dynamic 
behaviour of biological networks. A novel modelling tool, ANIMO, allows 
effective use of this approach and enables an intuitive construction of formal models.

ANIMO is not the first modelling tool to provide an interface to a
modelling formalism. Such interfaces exist in many other tools (see Suppl. Tab.~\ref{tab:tool-comparison}). With its
focus on user-friendliness and intuitive modelling, ANIMO's main contribution lies 
in making computational modelling more accessible to experts in biology.
Making use of the visual
interface provided by Cytoscape, network representations subscribe to biological conventions. 
Model parameters are kept to a minimum and can be directly accessed by mouse-clicking on 
nodes and edges. Because of the automatic translation of the network topology and 
user-defined parameters into an underlying formal model, training in the use of formal methods 
is not needed. In Supplementary Section~\ref{suppl:comparison-table}, a more in-depth
comparison between ANIMO and other modeling tools is given. For this comparison we selected a tool
for each of the most commonly used formalisms, and used criteria with a strong focus on 
user-friendliness.

In Section~\ref{sec:animo-drosophila}, we described the construction of an ANIMO
model of the ciracadian clock in \emph{Drosophila Melanogaster}. This model
captured the dynamics of the regualtory network and led to similar 
conclusions as an ODE model that had been
published previously~\cite{drosophila-ode-model}. This finding supports the use of
the series of modelling abstractions that we proposed. The biggest
difference between the construction of these models is that the model by~\cite{drosophila-ode-model}
is constructed by writing a system of mathematical equations, together
with an algorithm for simulation. In ANIMO, instead, a number of network
nodes is drawn for the molecules involved. 
These nodes are then linked by directed
interactions that represent cause-and-effect relationships, with a single parameter 
that defines the strength of each
interaction. This is a more intuitive approach to construct a model.
Further contributing to an interactive modelling process
is the compositionality of the model. Each node in the network
can be disabled at any time by the user, or extra nodes can be added,
without having to change any of the existing interactions.

In Section~\ref{subsec:case-study-larger}, we showed the construction of an executable model
of signalling events downstream of
TNF$\alpha$ and EGF in human colon carcinoma cells. This data set has been used for
previous modelling studies, based on partial least-squares regression and fuzzy 
logic~\cite{pathway-leastsquare,pathway-fuzzy}.
The partial least-squares regression model describes an abstract data-driven model 
that uses statistical correlations
to relate signal transduction events to various cellular decisions. This type of modelling is
very useful in uncovering new and unexpected relations. It is also successful in making
predictions, but gives little direct in the dynamic behaviour of the network. Fuzzy
logic analysis led to a model that gave a better fit to the dynamic network behaviour than
discrete logic (Boolean) models. Inspection of the inputs to the logical gates that were used
to model protein behaviour led to the prediction of novel interactions between proteins,
showing the usefulness of this approach. For most of the proteins, such as JNK1, time was
used as an input parameter. For example, the logical gates ``if TNF$\alpha$ is high
\emph{AND} time is low, then JNK1 is high'' and ``if TNF$\alpha$ is high \emph{AND} time is
high, then JNK1 is low'' were used to
describe the dynamic behaviour of JNK1. Although this leads to a representative
description of the dynamic behaviour of JNK1, peaks in protein activity at early time points were
not reproduced by the fuzzy logic model. Moreover, it gives no insight in the molecular interactions 
that are involved in activation or inhibition.

In this study, we used the same data set and performed a single round of the empirical cycle. This cycle starts off with the experiments carried out by~\cite{pathway-compendium}. We used the resulting experimental data, together with knowledge from curated databases~\cite{kegg,phosphosite} to construct an executable model of the biological system.
In contrast with the two approaches described above, ANIMO is aimed at the construction of
more mechanistic models, mimicking biochemical interactions \emph{in silico}. This way of modelling
gives a different type of insight. In the process of model construction, we extended a
prior-knowledge network with time-dependent extracellular crosstalk that has been reported
previously~\cite{pathway-autocrine}. To come up with possible explanations for a disagreement
between the model and the experimental data, two additional layers of
crosstalk were introduced, at the signal transduction and transcriptional level. These modifications 
improved the fit of the model to the data and can be interpreted as novel testable hypotheses.
Finally, we proposed new experiments that could be carried out to test these hypotheses, closing the empirical cycle. 
Together, our model sheds more light on the intricate
entanglement between the TNF$\alpha$ and EGF pathways at multiple cellular levels.
But above all,  the model provides an excellent starting point for further investigation.
Every new round in the empirical cycle will lift the understanding of the system to a higher level, leading to an incremental build-up of knowledge and an upward empirical spiral. Being intuitively accessible, ANIMO models facilitate sharing knowledge within and between groups and encourage collaborations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
The authors declare that they have no competing interests.

\section*{Author's contributions}
S.S. designed and performed the experiments, developed the Cytoscape integration, wrote
the manuscript; J.S. conceived, designed and performed the experiments, wrote the
manuscript; B.W. developed the Cytoscape integration, and designed and developed the
time slider UI; P.E.vdV. initiated the study, conceived the Cytoscape implementation, supervised the
project; R.A.U.C. performed experiments; M.K. designed experiments, analyzed data and
wrote the manuscript; R.L. Contributed to the methodology and supervised the project;
J.vdP. contributed to the strategy and methodology in the manuscript, in particular the
connection with formal methods; J.N.P. designed experiments, analyzed data sets, contributed
in particular to the application of ANIMO for large biological data, wrote the manuscript,
and supervised the project.

\section*{Acknowledgements}
We would like to thank Christof Francke for valuable discussions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
%\bibliographystyle{bmc-mathphys} % Style BST file
%\bibliography{Paper}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

% \bibliographystyle{bmc-mathphys}
% \bibliography{Paper}

\begin{thebibliography}{}

\bibitem[Aldridge et~al., 2009]{pathway-fuzzy}
Aldridge, B.~B., Saez-Rodriguez, J., Muhlich, J.~L., Sorger, P.~K., and
  Lauffenburger, D.~A. (2009).
\newblock {Fuzzy logic analysis of kinase pathway crosstalk in
  TNF/EGF/Insulin-induced signaling}.
\newblock {\em PLoS Comput Biol}, 5(4):e1000340+.

\bibitem[Alur and Dill, 1994]{timed-automata-alur-dill}
Alur, R. and Dill, D.~L. (1994).
\newblock A theory of timed automata.
\newblock {\em Theor. Comput. Sci.}, 126(2):183--235.

\bibitem[Bailly et~al., 1996]{ap1-il1a}
Bailly, S., Fay, M., Isra{\"{e}}l, N., and Gougerot-Pocidalo, M.~A. (1996).
\newblock The transcription factor {AP-1} binds to the human interleukin 1
  alpha promoter.
\newblock {\em European Cytokine Network}, 7(2):125--128.

\bibitem[Bannister et~al., 1994]{cfos-cjun}
Bannister, A.~J., Brown, H.~J., Sutherland, J.~A., and Kouzarides, T. (1994).
\newblock {Phosphorylation of the c-Fos and c-Jun HOB1 motif stimulates its
  activation capacity.}
\newblock {\em Nucleic Acids Res}, 22(24):5173--6+.

\bibitem[Bartocci et~al., 2009]{bartocci-oscillators}
Bartocci, E., Corradini, F., Merelli, E., and Tesei, L. (2009).
\newblock Model checking biological oscillators.
\newblock {\em Electronic Notes in Theoretical Computer Science},
  229(1):41--58.
\newblock Proceedings of the Second Workshop From Biology to Concurrency and
  Back (FBTC 2008).

\bibitem[Batt et~al., 2007]{oded-ode-ta-discretization}
Batt, G., Salah, R.~B., and Maler, O. (2007).
\newblock On timed models of gene networks.
\newblock In {\em Proceedings of the 5th international conference on Formal
  modeling and analysis of timed systems}, FORMATS'07, pages 38--52, Berlin,
  Heidelberg. Springer-Verlag.

\bibitem[Bengtsson et~al., 2002]{ta-audio-protocol}
Bengtsson, J., Griffioen, W.~D., Kristoffersen, K.~J., Larsen, K.~G., Larsson,
  F., Pettersson, P., and Yi, W. (2002).
\newblock Automated verification of an audio-control protocol using {UPPAAL}.
\newblock {\em The Journal of Logic and Algebraic Programming}, 52-53(0):163 --
  181.

\bibitem[Bonzanni et~al., 2009]{petri-nets2}
Bonzanni, N., Krepska, E., Feenstra, K.~A., Fokkink, W., Kielmann, T., Bal,
  H.~E., and Heringa, J. (2009).
\newblock Executing multicellular differentiation: quantitative predictive
  modelling of c.elegans vulval development.
\newblock {\em Bioinformatics/computer Applications in The Biosciences},
  25:2049--2056.

\bibitem[Ciocchetta and Hillston, 2009]{bio-pepa}
Ciocchetta, F. and Hillston, J. (2009).
\newblock {Bio-PEPA: A framework for the modelling and analysis of biological
  systems}.
\newblock {\em Theor. Comput. Sci.}, 410:3065--3084.

\bibitem[Davis, 2000]{jnk-signaling}
Davis, R.~J. (2000).
\newblock Signal transduction by the {JNK} group of {MAP} kinases.
\newblock {\em Cell}, 103(2):239--252.

\bibitem[{de Jong}, 2002]{hidde-review}
{de Jong}, H. (2002).
\newblock Modeling and simulation of genetic regulatory systems: A literature
  review.
\newblock {\em Journal of Computational Biology}, 9:67--103.

\bibitem[Dematt\'{e} et~al., 2008]{blenx}
Dematt\'{e}, L., Priami, C., and Romanel, A. (2008).
\newblock {Modelling and simulation of biological processes in BlenX}.
\newblock {\em SIGMETRICS Perform. Eval. Rev.}, 35:32--39.

\bibitem[D{\'{e}}rijard et~al., 1994]{ras-jnk1}
D{\'{e}}rijard, B., Hibi, M., Wu, I.-H., Barrett, T., Su, B., Deng, T., Karin,
  M., and Davis, R.~J. (1994).
\newblock {JNK1}: A protein kinase stimulated by uv light and ha-ras that binds
  and phosphorylates the c-{Jun} activation domain.
\newblock {\em Cell}, 76(6):1025--1037.

\bibitem[Efroni et~al., 2003]{interacting-sm2}
Efroni, S., Harel, D., and Cohen, I.~R. (2003).
\newblock Toward rigorous comprehension of biological complexity: Modeling,
  execution, and visualization of thymic t-cell maturation.
\newblock {\em Genome Research}, 13(11):2485--2497.

\bibitem[Fathallah-Shaykh et~al., 2009]{drosophila-ode-model}
Fathallah-Shaykh, H.~M., Bona, J.~L., and Kadener, S. (2009).
\newblock Mathematical model of the drosophila circadian clock: Loop regulation
  and transcriptional integration.
\newblock {\em Biophysical Journal}, 97(9):2399 -- 2408.

\bibitem[Fisher et~al., 2005]{interacting-sm1}
Fisher, J., Piterman, N., Hubbard, E. J.~A., Stern, M.~J., and Harel, D.
  (2005).
\newblock Computational insights into caenorhabditis elegans vulval
  development.
\newblock {\em Proceedings of the National Academy of Sciences of the United
  States of America}, 102(6):1951--1956.

\bibitem[Gaudet et~al., 2005]{pathway-compendium}
Gaudet, S., Janes, K.~A., Albeck, J.~G., Pace, E.~A., Lauffenburger, D.~A., and
  Sorger, P.~K. (2005).
\newblock {A compendium of signals and responses triggered by prodeath and
  prosurvival cytokines.}
\newblock {\em Mol Cell Proteomics}, 4(10):1569--1590.

\bibitem[Hessel and Pettersson, 2007]{ta-wap-gateway}
Hessel, A. and Pettersson, P. (2007).
\newblock Model-based testing of a wap gateway: an industrial case-study.
\newblock In {\em Proceedings of the 11th international workshop, FMICS 2006
  and 5th international workshop, PDMC conference on Formal methods:
  Applications and technology}, FMICS'06/PDMC'06, pages 116--131, Berlin,
  Heidelberg. Springer-Verlag.

\bibitem[Hornbeck et~al., 2004]{phosphosite}
Hornbeck, P.~V., Chabra, I., Kornhauser, J.~M., Skrzypek, E., and Zhang, B.
  (2004).
\newblock {PhosphoSite}: A bioinformatics resource dedicated to physiological
  protein phosphorylation.
\newblock {\em Proteomics}, 4(6):1551--1561.

\bibitem[Janes et~al., 2005]{pathway-leastsquare}
Janes, K.~A., Albeck, J.~G., Gaudet, S., Sorger, P.~K., Lauffenburger, D.~A.,
  and Yaffe, M.~B. (2005).
\newblock A systems model of signaling identifies a molecular basis set for
  cytokine-induced apoptosis.
\newblock {\em Science}, 310(5754):1646--1653.

\bibitem[Janes et~al., 2006]{pathway-autocrine}
Janes, K.~A., Gaudet, S., Albeck, J.~G., Nielsen, U.~B., Lauffenburger, D.~A.,
  and Sorger, P.~K. (2006).
\newblock The response of human epithelial cells to {TNF} involves an inducible
  autocrine cascade.
\newblock {\em Cell}, 124(6):1225--1239.

\bibitem[Kanehisa and Goto, 2000]{kegg}
Kanehisa, M. and Goto, S. (2000).
\newblock {KEGG: Kyoto Encyclopedia of Genes and Genomes}.
\newblock {\em Nucleic Acids Research}, 28(1):27--30.

\bibitem[Killcoyne et~al., 2009]{cytoscape}
Killcoyne, S., Carter, G.~W., Smith, J., and Boyle, J. (2009).
\newblock {Cytoscape: a community-based framework for network modeling.}
\newblock {\em Methods in molecular biology (Clifton, N.J.)}, 563:219--239.

\bibitem[Larsen et~al., 1997]{uppaal}
Larsen, K.~G., Pettersson, P., and Yi, W. (1997).
\newblock {UPPAAL} in a nutshell.
\newblock {\em International Journal on Software Tools for Technology Transfer
  (STTT)}, 1:134--152.

\bibitem[Man et~al., 2011]{ta-radiazioni}
Man, K., Krilavicius, T., Wan, K., Hughes, D., and Lee, K. (2011).
\newblock Modeling and analysis of radiation therapy system with respiratory
  compensation using {UPPAAL}.
\newblock In {\em Proceedings of the 9th IEEE International Symposium on
  Parallel and Distributed Processing with Application - {ISPA} 2011}, pages
  50--54.

\bibitem[{Marques Jr.} et~al., 2013]{ta-ws-bap}
{Marques Jr.}, A.~P., Ravn, A.~P., Srba, J., and Vighio, S. (2013).
\newblock Model-checking web services business activity protocols.
\newblock {\em International Journal on Software Tools for Technology
  Transfer}, 15(2):125--147.

\bibitem[Mendoza et~al., 1999]{boolean-networks-flower}
Mendoza, L., Thieffry, D., and Alvarez-Buylla, E.~R. (1999).
\newblock Genetic control of flower morphogenesis in arabidopsis thaliana: a
  logical analysis.
\newblock {\em Bioinformatics}, 15(7):593--606.

\bibitem[Nakano and Yamaguchi, 2011]{ta-giapponesi}
Nakano, S. and Yamaguchi, S. (2011).
\newblock Two modeling methods for signaling pathways with multiple signals
  using {UPPAAL}.
\newblock In {\em Proceedings of the 2nd International Workshop on Biological
  Processes \& Petri Nets {(BioPPN2011)}}, pages 87--101. CEUR-WS.org.

\bibitem[Reisig, 2011]{petri-nets}
Reisig, W. (2011).
\newblock Petri nets.
\newblock In Koch, I., Reisig, W., Schreiber, F., Dress, A., Vingron, M.,
  Myers, G., Giegerich, R., Fitch, W., and Pevzner, P.~A., editors, {\em
  Modeling in Systems Biology}, volume~16 of {\em Computational Biology}, pages
  37--56. Springer London.
\newblock 10.1007/978-1-84996-474-6\_3.

\bibitem[Russell et~al., 1995]{ras-mekk1}
Russell, M., Lange-Carter, C.~A., and Johnson, G.~L. (1995).
\newblock Direct interaction between {Ras} and the kinase domain of
  mitogen-activated protein kinase kinase kinase ({MEKK1}).
\newblock {\em Journal of Biological Chemistry}, 270(20):11757--11760.

\bibitem[Schivo et~al., 2012]{animo-bibe}
Schivo, S., Scholma, J., Wanders, B., {Urquidi Camacho}, R.~A., van~der Vet,
  P.~E., Karperien, M., Langerak, R., van~de Pol, J., and Post, J.~N. (2012).
\newblock Modelling biological pathway dynamics with {Timed Automata}.
\newblock In {\em 2012 IEEE 12th International Conference on Bioinformatics \&
  Bioengineering (BIBE)}, pages 447--453, Los Alamitos, CA, USA. IEEE Computer
  Society.

\bibitem[Shmulevich and Zhang, 2002]{boolean-networks2}
Shmulevich, I. and Zhang, W. (2002).
\newblock {Binary analysis and optimization-based normalization of gene
  expression data}.
\newblock {\em Bioinformatics}, 18(4):555--565.

\bibitem[Siebert and Bockmayr, 2008]{ta-siebert}
Siebert, H. and Bockmayr, A. (2008).
\newblock Temporal constraints in the logical analysis of regulatory networks.
\newblock {\em Theor. Comput. Sci.}, 391(3):258--275.

\bibitem[Thomas, 1973]{thomas-formalism}
Thomas, R. (1973).
\newblock Boolean formalization of genetic control circuits.
\newblock {\em J. Theor. Biol.}, 42(3):563--585.

\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}
  \begin{figure}[h!]
  \caption{\csentence{Sample figure title.}
      A short description of the figure content
      should go here.}
      \end{figure}

\begin{figure}[h!]
  \caption{\csentence{Sample figure title.}
      Figure legend text.}
      \end{figure}

\begin{figure}[!htb]
  \centering
  \includegraphics[width=0.35\textwidth]{images/empirical_spiral4}
 \caption{The empirical spiral: applying the empirical cycle in successive
rounds leads to a gradual build-up of knowledge.}\label{fig:empirical-spiral}
\end{figure}

\def\mekTA{\includegraphics[scale=.098]{images/abstraction_ta_erk3}}
\newlength\mekTAheight
\setlength\mekTAheight{\heightof{\mekTA}}
\begin{figure}[!hb]
%\begin{minipage}{\textwidth}
\begin{center}
\subfloat[\label{subfig:mek-erk}]{\begin{minipage}[c][\mekTAheight]{0.13\textwidth}\begin{center}\includegraphics[scale=.098]{images/abstraction_ta_mek-erk3}\end{center}\end{minipage}}
\qquad
\subfloat[\label{subfig:mek}]{\begin{minipage}[c][\mekTAheight]{0.13\textwidth}\begin{center}\includegraphics[scale=.098]{images/abstraction_ta_mek}\end{center}\end{minipage}}
\qquad
\subfloat[\label{subfig:erk}]{\begin{minipage}[c][\mekTAheight]{0.14\textwidth}\begin{center}\mekTA\end{center}\end{minipage}}
\end{center}
\caption{Formalization of an activation interaction into a \tas\ model.
{\bf \protect\subref{subfig:mek-erk}}~Classical depiction of a well-studied intracellular signal transduction 
interaction: MEK activates downstream protein ERK.
{\bf \protect\subref{subfig:mek}}~A \ta\ model of MEK, consisting of a single location (circle, active\_MEK) and 
a single transition (arrow). In this example, MEK activity is not regulated and MEK is always active. ${\sf t} < 20$ 
is termed an invariant on the location, allowing residence in this location as long as local
clock time {\sf t} is smaller than $20$ units. ${\sf t} > 18$ is termed a guard on the transition, allowing the
transition to take place when local clock {\sf t} is greater than $18$ units. Together, the invariant and guard
ensure that the transition must take place within the continuous time interval $18 < {\sf t} < 20$. When the
transition takes place, the action {\sf activate\_ERK!} is performed and the local clock is reset, ${\sf t} := 0$.
{\bf \protect\subref{subfig:erk}}~A \ta\ model of ERK, consisting of three locations, {\sf inactive\_ERK}
(the starting location, indicated by a double circle), {\sf 50\%\_active\_ERK} and {\sf 100\%\_active\_ERK},
and two transitions between the locations. Here, ERK has three activity levels: completely inactive, halfway active
and completely active.
A transition will take place when it is possible to synchronize with
the corresponding action {\sf activate\_ERK!} in the MEK automaton.
Each synchronization on channel {\sf activate\_ERK} represents the occurrence of the activating
interaction between MEK and ERK, and allows ERK to eventually become completely active. If we 
replace the time constraints for the occurrence of {\sf activate\_ERK!} with variables depending
on scenario~1 ($R = k \times [MEK]$), the second activation step would have the same time 
constraints as the first activation step, since the interaction rate only depends on MEK. If we use scenario~2
($R = k \times [MEK] \times [1 - ERK]$) instead, the time constraints are doubled after the first activation step, because only
50 \% of inactive ERK is left. The second activation step would then take twice the time of the first step.
}\label{fig:abstraction-mek-erk}
%\end{minipage}
\end{figure}

\def\modelGraphScale{0.2}%0.148}%0.215}
\def\legendGraphScale{0.23}%0.16}
\def\halfGraphScale{0.09}%0.067}%0.1075}
\begin{figure}[!bhtp]
\centering
\begin{tabular}{ll}
\subfloat{\includegraphics[scale=\legendGraphScale]{images/small-model-1g_legenda}}\addtocounter{subfigure}{-1}\subfloat[\label{fig:small-model-first}]{\includegraphics[scale=\modelGraphScale]{images/00-paper-model1f}}
& \subfloat[\label{fig:small-model-first-graph}]{\includegraphics[scale=\halfGraphScale]{images/00-paper-graph1m_riga}} \\[5ex]
\subfloat[\label{fig:small-model-third}]{\qquad\includegraphics[scale=\modelGraphScale]{images/00-paper-model3f}}
& \subfloat[\label{fig:small-model-third-graph}]{\includegraphics[scale=\halfGraphScale]{images/00-paper-graph3n_riga}} \\[5ex]
\subfloat[\label{fig:small-model-fourth}]{\includegraphics[scale=\modelGraphScale]{images/00-paper-model4g}}
& \subfloat[\label{fig:small-model-fourth-graph}]{\includegraphics[scale=\halfGraphScale]{images/00-paper-graph4o_riga}}
\end{tabular}
  \caption{
Incremental construction of an ANIMO model of signal transduction
events in human colon carcinoma cells upon stimulation with 100 ng/ml TNF$\alpha$.
Each construction step (top to bottom) is simulated in ANIMO, giving intermediate feedback
useful for the piecewise refinement of the model.
The graphs on the right show the dynamic behaviour of the corresponding models on the left, comparing it to the measured
activity values by~\cite{pathway-compendium} (error bars represent the standard deviation).
On the vertical axis, ``100'' represents the maximum protein activity in the complete experiment.
The red vertical line in each graph indicates a selected time point in the time course. 
Nodes in the corresponding network representation are coloured according to their activity at that time point.
All images in this figure are taken from the ANIMO user interface.
{\bf (\protect\subref*{fig:small-model-first}, \protect\subref*{fig:small-model-first-graph})}~Basic model showing direct activation of JNK1 and MK2 by TNF$\alpha$.
No peak dynamics are observed because no inactivating processes are present.
{\bf (\protect\subref*{fig:small-model-third}, \protect\subref*{fig:small-model-third-graph})}~The model after addition of inactivating phosphatases and a
negative feedback loop that down-regulates TNFR. Note that adding TNFR internalization or phosphatases alone would not be enough to reproduce activity peaks.
{\bf (\protect\subref*{fig:small-model-fourth}, \protect\subref*{fig:small-model-fourth-graph})}~The model after addition of IKK, IL1-secretion (abstracting
the autocrine IL-1 signalling described by~\cite{pathway-autocrine}), Casp8 and Casp3, showing the late response to TNF$\alpha$ signalling.\\
The {\sf \_{}data} suffix identifies experimental data; all other series are computed by ANIMO.\\
As the data set did not contain values for cleaved caspase-3, but only for its non-cleaved precursor pro-caspase-3,
we computed the {\sf Casp3\_{}data} series as $100\% - [\mbox{\sf pro-Casp3}]$.}\label{fig:small-model}
\end{figure}

\def\drosophilaGraphScale{0.069}%0.123}
\begin{figure}[!htpb]
\begin{center}
\includegraphics[width=0.49\textwidth]{images/drosophila_model5}%0.85\textwidth
% \subfloat[\label{subfig:drosophila-mrna}]{\includegraphics[scale=\drosophilaGraphScale]{images/drosophila_graph2}}\
% \subfloat[\label{subfig:drosophila-proteins}]{\includegraphics[scale=\drosophilaGraphScale]{images/drosophila_graph1}}
\end{center}
\caption{ANIMO model of the circadian clock in \emph{Drosophila Melanogaster} 
Autoregulatory negative feedback loops are present on each of the nodes of
the network, following the original model by~\cite{drosophila-ode-model}. These feedback loops ensure that
protein levels decrease over time when activating inputs are absent. The feedback loops are not represented here
for cosmetic reasons and clarity.
% \\ ANIMO plots of the concentration of
% mRNA~{\bfseries \protect\subref{subfig:drosophila-mrna}}
% and proteins~{\bfseries \protect\subref{subfig:drosophila-proteins}} over a period of 48~hours.\\
Naming conventions follow the same rules
as in the original model, with lower-case names representing mRNA, and upper-case names indicating proteins.
}\label{fig:drosophila-model}
\end{figure}

\def\largeModelScale{0.18}%0.15}%0.155}%0.27}
\def\legendScalaColori{0.21}%0.18}%0.21}
\def\legendScalaForme{0.21}%0.18}%0.21}
\def\scalaGrafici{0.0709}%0.13}
\begin{figure}[!htpb]
\centering
 \subfloat[\label{fig:large-model-no-hypotheses}]{\includegraphics[scale=\largeModelScale]{images/large_network_merged_no_hypotheses}}\\
%\subfloat{\includegraphics[scale=\legendScalaColori]{images/legenda_forme}}\\ \subfloat{\includegraphics[scale=\legendScalaForme]{images/legenda_colori}}\\
\subfloat{\includegraphics[scale=\legendScalaColori]{images/legenda_forme_e_colori}}\\
\addtocounter{subfigure}{-1}
\subfloat[\label{fig:large-model-graph1}]{\includegraphics[scale=\scalaGrafici]{images/TGF100_TNF0_ho_hyp_IL-1a3}}
\subfloat[\label{fig:large-model-graph3}]{\includegraphics[scale=\scalaGrafici]{images/TNF5_C225_no_hyp_JNK1_MK2}}
\caption{Signalling network downstream of TNF$\alpha$ and EGF in human colon carcinoma cells.
{\bf \protect\subref{fig:large-model-no-hypotheses}}
The model for the merged TNF$\alpha$ and EGF pathways. Node colours represent the
activity level of the corresponding modelled reactants at time $t = 15$ minutes after
a stimulation of 100 ng/ml TNF$\alpha$ + 100 ng/ml EGF.
{\bf \protect\subref{fig:large-model-graph1}}~Modelled production of IL-1$\alpha$ after stimulation with 100 ng/ml TGF$\alpha$ (24 hours).
{\bf \protect\subref{fig:large-model-graph3}}~Modelled activation of JNK1 and MK2 after stimulation with 5 ng/ml TNF$\alpha$ + 10 $\mu$g/ml C225 (2 hours).
\\
The {\sf \_{}data} suffix identifies experimental data; all other series are computed by ANIMO.}\label{fig:large-model-all}
\end{figure}

\begin{figure}[!tpb]
\begin{center}
\subfloat[\label{fig:large-model-complete}]{\includegraphics[scale=\largeModelScale]{images/large_network_legendg}}\\
 \subfloat[\label{fig:large-model-graph2}]{\includegraphics[scale=\scalaGrafici]{images/TGF100_vs_TNF100_hyp1_IL-1a4}}
 \subfloat[\label{fig:large-model-graph4}]{\includegraphics[scale=\scalaGrafici]{images/TNF5_C225_hyp2_JNK1_MK2}}
\end{center}
\caption{\scriptsize
{\bf \protect\subref{fig:large-model-complete}} The model for the merged TNF$\alpha$ and EGF pathways
after addition of the two hypotheses (highlighted).
Hypothesis 1 assumes IL-1$\alpha$ expression to depend on AP-1 activity, which in turn requires
both c-Jun en c-Fos to be activated by JNK1 and ERK, respectively. Hypothesis 2 assumes RAS as an activator
of MEKK1. Node colours represent the activity levels $15$ minutes
after stimulation of 100 ng/ml TNF$\alpha$ + 100 ng/ml EGF.
{\bf \protect\subref{fig:large-model-graph2}}~After the addition of the first hypothesis (activation of IL-1$\alpha$ production depending both
on JNK1 and ERK): production of IL-1$\alpha$ after stimulation with 100 ng/ml TNF$\alpha$ (series {\sf IL-1a~(TNFa)})
compared with stimulation with 100 ng/ml TGF$\alpha$ (series {\sf IL-1a~(TGFa)}) (24 hours).\\
{\bf \protect\subref{fig:large-model-graph4}}~After the addition of the second hypothesis (activation of MEKK1 downstream of EGFR):
stimulation with 5 ng/ml TNF$\alpha$ + 10 $\mu$g/ml C225 (2 hours).
Suppl. Sect.~\ref{suppl:parameters-tnf-egf} explains how the dosage of 5 ng/ml TNF$\alpha$ was represented in the model.\\
The {\sf \_{}data} suffix identifies experimental data; all other series are computed by ANIMO.}\label{fig:large-model-graph}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go.}
      \begin{tabular}{cccc}
        \hline
           & B1  &B2   & B3\\ \hline
        A1 & 0.1 & 0.2 & 0.3\\
        A2 & ... & ..  & .\\
        A3 & ..  & .   & .\\ \hline
      \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.

\end{backmatter}

\end{document}

\appendix
\clearpage
\setcounter{figure}{0}
\setcounter{table}{0}
\setcounter{page}{1}
\onecolumn

% \setlength{\oddsidemargin}{1.875in}
% \setlength{\evensidemargin}{1.875in}
% \addtolength{\textwidth}{-3.75in}

\newlength\addedmarginsingle
\setlength\addedmarginsingle{0.875in}
\newlength\addedmargintotal
\setlength\addedmargintotal{2\addedmarginsingle}

\addtolength{\oddsidemargin}{\addedmarginsingle}
\addtolength{\evensidemargin}{\addedmarginsingle}
\addtolength{\textwidth}{-\addedmargintotal}
\addtolength{\linewidth}{-\addedmargintotal}
\addtolength{\hsize}{-\addedmargintotal}
\addtolength{\topmargin}{\addedmarginsingle}
\addtolength{\textheight}{-\addedmargintotal}
\addtolength{\vsize}{-\addedmargintotal}

\pagestyle{plain}

\clearpage
% \thispagestyle{empty}\ \

\thispagestyle{empty}
\ \\ \ \\ \ \\ \ \\ \ \\
\begin{center}
 {\Huge Bringing biological networks}\\ \ \\ {\Huge to life with ANIMO}\\ \ \\ \ \\
 {\huge Additional Materials}
\end{center}
\clearpage

\makeatletter

% Copied from the LaTeX sources
\def\addcontentsline#1#2#3{%
  \addtocontents{#1}{\protect\contentsline{#2}{#3}{\thepage}}}
\long\def\addtocontents#1#2{%
  \protected@write\@auxout
    {\let\label\@gobble \let\index\@gobble \let\glossary\@gobble}%
    {\string\@writefile{#1}{#2}}}
\titlecontents{section} % set formatting for \section -
                        % \subsection must be formatted separately
[2.em]                 % adjust left margin
{\bf}             % font formatting
{\contentslabel{2.em}} % section label and offset
{\hspace*{-2.em}}
{\titlerule*[1pc]{}\contentspage}
\titlecontents{subsection} % set formatting for \section -
                        % \subsection must be formatted separately
[4.5em]                 % adjust left margin
{\rmfamily}             % font formatting
{\contentslabel{2.5em}} % section label and offset
{\hspace*{2.5em}}
{\titlerule*[1pc]{.}\contentspage}
% Copied from article.cls
%\setcounter{tocdepth}{5}
\newcommand\tableofcontents{%
\begin{spacing}{1.3}
    \section*{\contentsname
        \@mkboth{%
           \MakeUppercase\contentsname}{\MakeUppercase\contentsname}}%
    \@starttoc{toc}%
\end{spacing}
}
% \newcommand*\l@paragraph{\@dottedtocline{4}{7.0em}{4.1em}}
% \newcommand*\l@subparagraph{\@dottedtocline{5}{10em}{5em}}
% \newcommand\listoffigures{%
%     \section*{\listfigurename}%
%       \@mkboth{\MakeUppercase\listfigurename}%
%               {\MakeUppercase\listfigurename}%
%     \@starttoc{lof}%
%     }
% \newcommand*\l@figure{\@dottedtocline{1}{1.5em}{2.3em}}
% \newcommand\listoftables{%
%     \section*{\listtablename}%
%       \@mkboth{%
%           \MakeUppercase\listtablename}%
%          {\MakeUppercase\listtablename}%
%     \@starttoc{lot}%
%     }
% \let\l@table\l@figure

\makeatother

\addtocontents{toc}{\protect\setcounter{tocdepth}{5}}
\thispagestyle{empty}
\tableofcontents
\clearpage

\setcounter{page}{1}
\setcounter{section}{0}

\renewcommand\figurename{Figure}
\renewcommand*\thefigure{S\arabic{figure}}
\renewcommand*\thetable{S\arabic{table}}

\def\ta{TA}
\def\tas{TA}

\section*{Requirements and installation}\label{sec:animo-installation}
In order to run ANIMO, a desktop or laptop computer is needed with the following software installed:
\begin{itemize}
  \item Java: see Section~\ref{sec:install-java}
  \item Cytoscape~\cite{cytoscape}: see Section~\ref{sec:install-cytoscape}
  \item UPPAAL~\cite{uppaal}: see Section~\ref{sec:install-uppaal}
\end{itemize}
The software to run Java-based programs is provided for free by Oracle. More information
is available on the \url{java.com} website.
Cytoscape is an open source project released under the terms of the GNU Lesser General Public License.
UPPAAL is developed by a collaboration by the universities of Uppsala (Sweden) and Aalborg (Denmark),
and is free for non-commercial applications in academia only.

All of these softwares work under Windows~(\winsymbol), Mac-OS~(\macsymbol) and all most common GNU/Linux~(\linuxsymbol) distributions.
If the requirements are already met, ANIMO can be directly installed following the instructions in Section~\ref{sec:install-animo}.\\
\emph{Note}: when required to type something, the text to input will be represented ``{\tt like this}'': the quotation marks
are not intended to be typed.\\
In case of problems accessing a web site with Microsoft Internet Explorer, we advise to try with a different web browser (such as Mozilla Firefox or
Google Chrome) or to update Internet Explorer.

\subsection*{Java}\label{sec:install-java}
\begin{enumerate}
\item\label{step:open-console} In order to check that Java is installed, open a console
\winmaclinux{Windows 7: press Windows button and type ``{\tt cmd}'', then press Return. Previous versions: in the Start
menu find \emph{All programs} $\rightarrow$ \emph{Accessories} $\rightarrow$ \emph{Command Prompt}.}%
{Go to \emph{Applications} $\rightarrow$ \emph{Utilities} $\rightarrow$ \emph{Terminal}.}%
{Under Gnome, press Alt-F2, type ``{\tt gnome-terminal}'', then press Return. Under KDE, press the KMenu button, type
``{\tt konsole}'' and click Konsole. Under Unity, press the home button (the one with the Ubuntu
logo:~\ubuntusymbol), type ``{\tt terminal}'' and click the Terminal icon.}
\item Type ``{\tt java}'' and press Return. If a brief error message like ``{\tt unknown command}'' is shown, Java needs to be
installed: please proceed to step~\ref{step:install-java}. Otherwise, please continue to Section~\ref{sec:install-cytoscape}.
\item\label{step:install-java}
\winmaclinux{Point your web browser to \url{java.com}.\\
Click \emph{Free Java Download}, then on \emph{Agree and Start Free Download}.\\
If asked for permission to run the installer, grant that permission.\\
After the download, double click the installer and install Java following the guided steps.\\
To check that Java has been correctly installed, a web page is automatically opened at
the end of the installation process. Click on \emph{Verify Java version}. If a \emph{Congratulations!} message is shown,
Java has been successfully installed. Otherwise, please try again or contact Java support.}%
{Go to \emph{Applications} $\rightarrow$ \emph{Utilities} $\rightarrow$ \emph{Java Preferences}.\\
If the Java Preferences window is shown, Java is already installed, otherwise the system will prompt you to install
it. Follow the instructions and Java will be correctly installed at the end of the procedure.}%
{If you run Ubuntu, open the Software centre, search for ``java'' and select \emph{OpenJDK Java 6 Runtime}. An
\emph{Install} button
will appear next to the name of the package: click that button and Java will be correctly installed.\\
If you run another distribution, use your package manager in a similar way. If you cannot find OpenJDK, there may be
the possibility to install \emph{Oracle Java Development Kit (JDK)} instead.}
\end{enumerate}

\subsection*{Cytoscape}\label{sec:install-cytoscape}
Cytoscape can be found at the address \url{www.cytoscape.org/download.php}: an automatic installer
program can be downloaded. Please note that you need to register and accept Cytoscape's terms of use
before being able to start the download.
Choose the latest 2.x version (at least 2.8.3, on the left column),
possibly using a platform specific installer. For Windows, you can choose \emph{64bit} only
if you know that your computer can run 64bit programs, otherwise it is safe to choose \emph{32bit}.
% \begin{enumerate}
% \item Point your browser to \url{www.cytoscape.org/download.html}.
% \item Insert the required data and accept the terms of use.
% \item Download the installer for the latest version, choosing the correct platform.
% \item Once the download is finished, find the downloaded file and double click to start the installation.
% \item Once the installation is complete, you should find a \emph{Cytoscape} menu item in your
% \emph{Applications}/\emph{Programs} menu.
% \end{enumerate}

\subsection*{UPPAAL}\label{sec:install-uppaal}
\begin{enumerate}
\item Point your browser to \url{www.uppaal.org}.\\
\emph{Note}: UPPAAL is free only for
academic use. Information and contacts for commercial licenses can be found on the web site.
\item Click the \emph{Download} link, and choose the latest \emph{development} version (at least 4.1) for your operating system.
\item Fill in the required contact information and click the \emph{Accept and download} button to download UPPAAL.\\
\emph{Note}: problems with the registration on UPPAAL website have been reported when using some versions of Microsoft Internet Explorer.
If the registration is unsuccessful, please consider updating Internet Explorer or changing your web browser.
\item Unzip the downloaded file to a known location: UPPAAL will be installed there.
\item\label{step:mac-install-uppaal} Complete UPPAAL installation. \macpc{Open the UPPAAL installation location in Finder,
drop the \emph{UPPAAL.App} icon in your \emph{Applications} folder,
and copy the \emph{verifyta} executable file to a known location. The installation of UPPAAL is complete:
go to Section~\ref{sec:install-animo}.}{Open a console (this was done in Sec.~\ref{sec:install-java}, step
number~\ref{step:open-console}), type ``\url{cd} \url{PATH_TO_THE_UPPAAL_DIRECTORY}'' and press Return;
\url{PATH_TO_THE_UPPAAL_DIRECTORY} is the path to the directory where you installed UPPAAL. It can be for example
``\url{c:\Users\myuser\Desktop\uppaal-4.1.15}'',
 or
``\url{/home/myuser/programs/uppaal-4.1.15}''.\\
\winsymbol\ \emph{Note}: some Windows users may have access only to specific partitions (D:, Z:,\dots): in that case,
please first change to the corresponding
drive letter where the downloaded file was extracted. For example: if UPPAAL is located in \url{d:\myuser\Programs\uppaal-4.1.15},
the two commands to be entered are\\
``\url{d:}''\\
``\url{cd} \url{\user\Programs\uppaal-4.1.15}''}
\item Type ``{\tt java -jar uppaal.jar}'' and press Return.
\item The license for UPPAAL will be automatically acquired, and the main window of UPPAAL user interface will appear:
you may now close that window.
\end{enumerate}

\subsection*{Installing ANIMO}\label{sec:install-animo}
\begin{enumerate}
\item ANIMO is {\bfseries free only for academic use}. For commercial licenses, please contact us.
\item Run Cytoscape.
\item Click the menu command \emph{Plugins} $\rightarrow$ \emph{Manage Plugins}: the \emph{Manage Plugins}
window will open.
\item Select the \emph{Settings tab} and press the \emph{Add} button.
\item Insert this \emph{Name}: ``{\tt ANIMO}'', and this \emph{URL}:
``\url{http://fmt.cs.utwente.nl/tools/animo/plugins.xml}''
(please note: the ``\url{http://}'' is required), then confirm with \emph{OK}.
\item From the \emph{Download Sites} list in the upper part of the \emph{Manage Plugins} window, select \emph{ANIMO}
(you may need to scroll down: it should appear after \emph{Cytoscape}).
\item The panel on the left shows a smaller list of plugins: under \emph{Available for Install} $\rightarrow$
\emph{Analysis},
select \emph{ANIMO v1.0.37}.
\item Click the \emph{Install} button. The ANIMO tool will be automatically downloaded and installed.
\item The tool will ask you to indicate the position of the \emph{verifyta} executable, which is the tool to verify Timed Automata models.
You can find it in the \emph{bin} (\emph{bin-Linux}, \emph{bin-Win32}, \dots\ depending on your operating system) directory inside the
UPPAAL installation directory, or where it was copied at step~\ref{step:mac-install-uppaal} for \macsymbol\ in Section~\ref{sec:install-uppaal}.
\item Click the \emph{Close} button to close the \emph{Manage Plugins} window.
\item ANIMO is correctly installed and ready to be used.
\end{enumerate}

\clearpage
\section*{ANIMO user's manual}\label{sec:animo-manual}
We will now present a step-by-step sequence to obtain an example model
with ANIMO, which will allow us to illustrate the main features of the tool.

\subsection*{Modelling a small network}\label{sec:modeling-network-example}
\begin{enumerate}
\item Run Cytoscape.
\item If Cytoscape is already running and there are open documents, please make sure that the current work is saved before proceeding.
\item From the \emph{File} menu, select \emph{New} $\rightarrow$ \emph{Session}. Answer positively to the question
``\emph{Current session (all networks/attributes) will be lost. Do you want to continue?}''.
\item From the \emph{File} menu, select \emph{New} $\rightarrow$ \emph{Network} $\rightarrow$ \emph{Empty Network}.
\item In the \emph{Control Panel} find the \emph{Editor} tab. If you cannot find it,
click the black arrows on the top right of the panel to search through the available tabs. Click the name of the tab to activate it.
\item\label{step:add-nodes} Add $5$ nodes to the empty network by Ctrl-clicking
on empty areas of the \emph{Network} window.\\
\emph{Note}: \emph{Ctrl-clicks} are obtained as follows. While
holding the {\tt Ctrl} key down, click with the left mouse button, then release the {\tt Ctrl} key.
The {\tt Ctrl} key is usually located in the lower left or lower right corner of the keyboard. Apple keyboards
may have the \maccmd\ symbol instead of {\tt Ctrl}.
\item The \emph{Edit reactant} dialogue window is opened when a new node is added,
or when you right click an existing node and then select the \emph{[ANIMO] Edit reactant\dots}
item from the menu. Use that window to set the properties of the nodes as indicated in Table~\ref{tab:setting-nodes},
taking the setting in Figure~\ref{fig:edit-reactant} for node A as reference. When
the properties of a node have been inserted, confirm the choice with the \emph{Save} button.
\newcounter{miocounterperenumerate}
\setcounter{miocounterperenumerate}{\value{enumi}}
\end{enumerate}\vspace{-2ex}

\begin{table}[htbp]
\begin{minipage}{\textwidth}
\processtable{The settings for the nodes (signalling network components) in the example.\label{tab:setting-nodes}}
{\begin{tabular}{llllll}%|c|c|c|c|c|c|}
\ \\
\hline\noalign{\vskip 2mm}
  \multirow{2}{*}{{\bfseries Name}} & {\bfseries Total act.} & {\bfseries Initial act.} & \multirow{2}{*}{{\bfseries Molecule type}} &
\multirow{2}{*}{{\bfseries Enabled?}} & \multirow{2}{*}{{\bfseries Plotted?}}\\
& {\bfseries levels} & {\bfseries level} & & & \\[2mm]
\hline\noalign{\vskip 2mm}
  A & 15 & 15 & Cytokine & Yes & No\\[5mm]
  B & 15 & 0 & Receptor & Yes & No\\[5mm]
  C & 15 & 0 & Other & Yes & No\\[5mm]
  D & 100 & 0 & Kinase & Yes & Yes\\[5mm]
  E & 1 & 1 & Phosphatase & Yes & No\\[2mm]
\hline
\end{tabular}
}{}
\end{minipage}
\end{table}

\begin{figure}[htpb]
\begin{minipage}{\textwidth}
\begin{center}
 \includegraphics[width=.5\textwidth]{images/edit_reactantA}\\
 \caption{The \emph{Edit reactant} window: modifying the properties of node A.}\label{fig:edit-reactant}
\end{center}
\end{minipage}
\end{figure}

\begin{enumerate}
\setcounter{enumi}{\value{miocounterperenumerate}}
\item\label{step:add-edges} In order to add edges to the network, make sure that the \emph{Editor} tab is still active
in the \emph{Control Panel}, and
add the following edges by Ctrl-clicking the source and then clicking the target: A~$\rightarrow$~B, B~$\rightarrow$~C,
C~$\rightarrow$~B, B~$\rightarrow$~D, E~$\rightarrow$~D.
\item The \emph{Edit reaction} dialogue window is opened when you add a new edge,
or when you right click an existing edge and then select the \emph{[ANIMO] Edit reaction\dots} item
from the menu. Use that window to set the parameters of the edges as indicated in Table~\ref{tab:setting-edges}. The settings
for the edge A $\rightarrow$ B should reflect the ones shown in the \emph{Edit reaction} window in Figure~\ref{fig:edit-reaction}.\\
\emph{Note}: In order to insert a qualitative parameter like the ones required by the example network,
click once the slider in the \emph{parameter} box to activate it, and then move the slider to match the requested value.
\setcounter{miocounterperenumerate}{\value{enumi}}
\end{enumerate}

\begin{table}[!ht]
\begin{minipage}{\textwidth}
\processtable{The settings for the edges (interactions) in the example.\label{tab:setting-edges}}
{\begin{tabular}{llll}%|c|c|c|c|}
\ \\
\hline\noalign{\vskip 2mm}
  {\bfseries Interaction} & {\bfseries Influence} & {\bfseries Scenario} & {\bfseries Parameter value}\\[2mm]
\hline
\noalign{\vskip 2mm}  A $\rightarrow$ B & Activation & 1 & Medium \\[5mm]
\noalign{\vskip 2mm}  B $\rightarrow$ C & Activation & 1 & Slow \\[5mm]
\noalign{\vskip 2mm}  C $\rightarrow$ B & Inhibition & 1 & Fast \\[5mm]
\noalign{\vskip 2mm}  B $\rightarrow$ D & Activation & 1 & Slow \\[5mm]
\noalign{\vskip 2mm}  E $\rightarrow$ D & Inhibition & 2 & V. Slow \\[2mm]
\hline
\end{tabular}}{}
\end{minipage}
\end{table}\vspace{-2ex}

\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\begin{center}
\includegraphics[width=.5\textwidth]{images/edit_reactionAB}\\
\caption{The \emph{Edit reaction} window: modifying the properties of edge A $\rightarrow$ B.}\label{fig:edit-reaction}
\end{center}
\end{minipage}
\end{figure}

\begin{enumerate}
\setcounter{enumi}{\value{miocounterperenumerate}}
\item In the \emph{Control Panel} activate the \emph{ANIMO} tab by clicking its title.
\item Click the \emph{Choose seconds/step} button. A new dialogue window will appear: you can
safely choose a time resolution of 1 second per step and click \emph{OK}.
\item Click the \emph{Analyse network} button.
\item After a few seconds the \emph{Results Panel} should appear on the right,
showing a plot of the activity level of reactant D over a time course of 120 minutes.
Figure~\ref{fig:rete-esempio} shows the resulting network and graph plot.
\end{enumerate}

\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\begin{center}
  \includegraphics[width=.9\textheight, angle=90]{images/esempio_uso_ANIMO3}
\end{center}
\caption{The completed example, where also the feature that allows to view the activity levels
of reactants at chosen simulation times is demonstrated: the vertical red bar in the graph on the
right can be moved through the slider under the graph, and indicates the point in the time series
on which the colouring of the nodes in the \emph{Network} window is based.
The legends for colours and shapes can be found in the \emph{Legend} panel.}\label{fig:rete-esempio}
\end{minipage}
\end{figure}

\subsubsection*{Managing simulation data and activity levels plots}
Each time a simulation result is obtained, a new tab is added to the \emph{Results Panel} (see the right part of Fig.~\ref{fig:rete-esempio})
in which we identify three buttons, a plot of the activity levels of the selected reactants and a time slider.

Clicking on the button \emph{Change title} allows to select a new title for the tab: this can be useful e.g. when comparing different
simulations made on similar configurations of the same network. Button \emph{Save simulation data\dots} allows to save
the simulation data of the current tab on a \emph{.sim} file, which can then be loaded and inspected in the future.
The \emph{Load simulation data\dots} button in the \emph{Control Panel} above the \emph{Simulation} box can be used for this purpose.
Please note that the best results are obtained only when loading a \emph{.sim} file when the \emph{Network} window contains the same network
on which the simulation data are based. If no network is currently opened, a network will \emph{not} be opened by loading a \emph{.sim} file.
The \emph{Close} button is used to close the currently displayed results tab.

Right clicking inside the graph area will bring up a menu that allows to perform some basic operations with the graph
and its data:
\begin{itemize}
  \item \emph{Add data from CSV}: \label{csv-import-format}superpose the graph with other data series found in a \emph{.csv} (comma separated values) file. This file type can be
obtained for example by exporting data from the default Excel format. If you want the data in the \emph{.csv} file to be rescaled so
that its maximum Y value coincides with the maximum in the plot, the data file needs to contain a column named (exactly)
\emph{Number\_{}of\_{}levels}, on the first row of which the maximum of the scale for the \emph{.csv} data needs to be put. For example,
if the data in the \emph{.csv} file are on a 0-100 scale, the value for \emph{Number\_{}of\_{}levels} will be 100.
  \item \emph{Save as PNG}: save the graph as it is shown in a \emph{.png} image file. This file format can be opened by most
image editors.
  \item \emph{Export visible as CSV}: export to a \emph{.csv} file all the series that are currently visible (i.e., not hidden)
in the graph.
  \item \emph{Clear Data}: clear the contents of the graph, removing all series. This can be useful for plotting
a \emph{.csv} file without superposing it to the current graph, or for loading a file in which all hidden data where removed
(exporting the visible graph to a \emph{.csv} with the previous command).
  \item \emph{Graph interval}: change the lower and upper bounds for X and Y axes.
  \item \emph{Zoom rectangle}: zoom the graph around a user-chosen rectangular area.
    After selecting this command the shape of the mouse cursor changes into a cross. The area of the plot to be
    zoomed can then be selected by dragging a rectangular selection around it (see the definition of \emph{rectangular
    selection} on page~\pageref{nota:rectangular-selection}).
  \item \emph{Zoom extents}: bring the zoom level back to default, cancelling the effects of any \emph{Zoom rectangle}
command.
\end{itemize}

Whenever the result of one or more simulations is shown as a graph, it is possible to use the slider under the graph to
move through the entire simulation, showing the activity levels of all reactants represented with different node colouring in
the \emph{Network} window on the left. For an example, see Figure~\ref{fig:rete-esempio}: the vertical red line in the
graph represents the time instant on which the colours of the nodes in the \emph{Network} window are based, and can
be moved with the slider over which the mouse cursor is drawn.

\subsection*{Additional tips}
\subsubsection*{Editing a network in Cytoscape}
Nodes and arcs can be placed in the network as shown previously: with the \emph{Editor} tab selected in the
\emph{Control Panel},
Ctrl-click (click while holding the {\tt Ctrl} or \maccmd\ key) in an empty place to add a node; Ctrl-click the source
node and click the target node to add an arc. It is also possible to drag and drop the node/arc icons from the \emph{Control Panel}
into the \emph{Network} window.\\
\emph{Note}: to perform \emph{drag and drop} move the mouse cursor over the icon in the \emph{Control Panel}, click with the left mouse
button and, without releasing the button, drag the mouse cursor on the \emph{Network} window where you want to
place the symbol; then release the left mouse button.

In order to delete a node/edge, select it by clicking
or grouping them in a larger rectangular selection, and then press the \emph{Delete} key on the keyboard
or select the \emph{Edit} $\rightarrow$ \emph{Delete Selected Nodes and Edges} menu command.\\
\emph{Note}: in order to obtain a \emph{rectangular selection}\label{nota:rectangular-selection},
left click in the \emph{Network} window where the upper-left corner of the rectangle should be and,
without releasing the left mouse button, drag the mouse cursor to where the lower-right corner of the
rectangular selection should be; then release the left mouse button. All the entities which were
\emph{even partially} touched by the rectangle are now selected.

Navigation inside the \emph{Network} window can be performed by clicking and dragging the centre mouse button,
while zooming can be done by either rotating the mouse wheel or clicking and holding the right mouse button
while moving the mouse in a vertical direction.

Finally, note that the colours used to represent node activity can be changed using the \emph{VizMapper}
interface provided by Cytoscape, changing the setting for \emph{Node colour}, shown in Figure~\ref{fig:change-gradient}.
To change the node colours, activate the \emph{VizMapper\texttrademark} tab in the \emph{Control Panel}, and
find the entry named \emph{Node Colour} in the \emph{Visual Mapping Browser} box. Click the arrow-shaped
icon directly to the left of \emph{Node Colour}: the current setting for the node colours should appear:
click the coloured bar to open the window shown in Figure~\ref{fig:change-gradient}.
The \emph{activityRatio} on which the node colouring is based is the ratio between the current activity level
of a node and its number of activity levels.
The coloured arrows pointing downward on the upper border of the coloured rectangle can be dragged along the length of the $[0, 1]$
interval, thus changing the point at which a particular colour appears. New arrows can be added with the \emph{Add} button, while
clicking an arrow and pressing the \emph{Delete} button will remove one. To change the colour of a point in the interval,
double click the corresponding arrow, and a new window will open, allowing you to choose a new colour: clicking \emph{OK} will
accept the new colour. The modifications made in the \emph{Gradient Editor for Node Colour} window should be automatically reflected in
the model: when the gradient is as wanted, simply close the window. If the node colours seem not to have been updated, please move the
slider under an existing graph in the \emph{Results Panel}.

\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\centering
  \includegraphics[width=.6\textwidth]{images/editing-gradient2}
  \caption{Changing the colours used to represent node activity in an ANIMO model.}\label{fig:change-gradient}
\end{minipage}
\end{figure}

\subsubsection*{ANIMO features}
Nodes and edges (and groups of nodes/edges highlighted via a rectangular selection)
can be disabled by choosing \emph{[ANIMO] Enable/disable} from
the right-click menu: they will be represented with less saturated colours and can be re-enabled by performing
the same action. Moreover, a node can be enabled/disabled directly in its properties window, where it is
also possible to add/remove the node from the list of series appearing in the graph resulting from a simulation
of the network by selecting \emph{Plotted} or \emph{Hidden} (see also Figure~\ref{fig:edit-reactant}):
nodes that will be plotted are circled in blue.
Every enabled node will be taken into account when computing the evolution of the system,
but only nodes marked as \emph{Plotted} will appear in a graph.

Each plot in an \emph{ANIMO Results} tab contains by default a legend, which can be used to modify which series are
displayed and how they are displayed. Clicking with the central mouse button on a series name will hide it from the
graph, while the same centre-click the coloured line beside the series name will change the colour of that series,
cycling through a predefined set of available colours. The entire legend can be hidden by clicking with the
central mouse button anywhere on the graph (not inside the legend), or it can be dragged around by clicking and holding the left
mouse button, releasing it when the preferred position is reached. Rotating the mouse wheel will allow the thickness of all
the graph lines, and the size of the text, to grow or decrease: this feature can be useful when the window containing
the graph is very large.

As the model is non-deterministic, i.e. its
evolution will not be exactly the same for every single simulation run, it is possible to ask ANIMO to perform
a number of simulation runs in a batch, plotting the averages of the activity levels over the runs together with a standard
deviation value, or showing a so-called \emph{overlay plot} where all runs are plotted over each other. The controls
that allow to ask for multiple simulation runs can be found in the \emph{Control Panel}, inside the \emph{Simulation} box.

Standard deviation may be represented in the graph: it is normally shown as vertical bars, but its aspect can be
cycled through five possibilities (vertical bars, shading, both bars and shading, bars and symbols, none) by right-clicking the
corresponding line in the legend. Symbols associated to a representation of standard deviation can be changed by Shift-right-clicking
(holding down the \emph{Shift} key, right click) on the corresponding line in the legend.
Standard deviation values can be obtained when asking for multiple simulations in the network
analysis, but they can also be present in a \emph{.csv} file, e.g. when the file contains averages of experimental data.
In a \emph{.csv} file, the column containing the standard deviation values for column \emph{A}
should be named \emph{A\_{}StdDev} for the program to recognize it and properly display the data series
with the associated error values.

\subsubsection*{Parameter settings}
The application of some basic strategies when setting the parameters for a network allows
the less experienced users to considerably shorten the modelling time.
First of all, it is important to proceed in a \emph{top-to-bottom} order, trying to match
a component to the corresponding data before inserting the components downstream thereof.
Second, when choosing the kinetic parameter for an interaction, we advise to first use the qualitative settings (very slow, slow, medium, fast, very fast):
this allows to define the relative speeds of the interactions as soon as possible,
leaving the more precise parameter setting procedure as a follow-up step. Finally, as can be seen from
the parameter settings of Section~\ref{sec:modeling-network-example}, in order to obtain a peak
behaviour it is particularly important that
a negative feedback is present (as an example, see the interactions involving B and C in Tab.~\ref{tab:setting-edges}),
and that the inactivating interaction in the loop is faster than the ones activating the target node.

A final note on the \emph{seconds/step} button. This button allows to define the time granularity
of the simulations, but it is not strictly necessary to choose a very precise value.
% As the time bounds in the \tas\ model are integers, rounding errors can make the model
% behave differently from what the parameters define.
If the current value for \emph{seconds/step}
is too high (or too low) to allow the network to be properly simulated, ANIMO will automatically choose (respectively)
the highest (lowest) value that still allows to avoid rounding problems. It will be possible to notice
such a change in the value of time scale when the number on the \emph{seconds/step} button changes.

\subsubsection*{Updating ANIMO}
To check whether a new version of ANIMO has been published, run Cytoscape and ask for an update of all
plug-ins via the menu command \emph{Plugins} $\rightarrow$ \emph{Update Plugins}. After some seconds
during which Cytoscape will contact all the providers of the installed plug-ins,
the system should report the list of updatable plug-ins.\\
\emph{Note}: a window with the message \emph{Attempting to connect to XYZ\dots} may appear and disappear multiple times:
it is the normal behaviour.

If an updated version of ANIMO is available, it will appear
under the category \emph{Updatable Plugins} $\rightarrow$
\emph{Analysis} $\rightarrow$ \emph{ANIMO v1.0.37}. If no plug-in can be updated, a message stating \emph{No
updates available for currently installed plug-ins.} will be shown.

\clearpage
\section*{Naming conventions}\label{suppl-sec:names}
Table~\ref{suppl-tab:names} explains the abbreviations used in the paper.

\begin{center}
\begin{longtable}{lll}
\caption{Explanation of the abbreviated names referring to molecular species
in the main text.}\label{suppl-tab:names} \\
%\begin{tabular}{lll}
\hline
\noalign{\vskip 2mm} {\bfseries Abbreviation} & {\bfseries Full name} \\[2mm]
\hline
\endfirsthead

\multicolumn{3}{c}%
{\scriptsize{\bfseries \tablename\ \thetable{}} \--\ continued from previous page} \\
\hline
\noalign{\vskip 2mm} {\bfseries Abbreviation} & {\bfseries Full name} \\[2mm]
\hline
\endhead

\hline \multicolumn{3}{r}{{Continued on next page}} \\ %\hline
\endfoot

\hline
\endlastfoot

\noalign{\vskip 2mm}  Akt & protein kinase B & P31749\\
\noalign{\vskip 2mm}  AP-1 & activator protein 1 & heterodimer \\
                            & & of c-Jun and c-Fos\\
\noalign{\vskip 2mm}  Casp3 & caspase 3 & P42574\\
\noalign{\vskip 2mm}  Casp8 & caspase 8 & Q14790\\
\noalign{\vskip 2mm}  c-Fos & proto-oncogene-protein c-fos & P01100\\
\noalign{\vskip 2mm}  c-Jun & Jun activation domain binding protein & P05412\\
\noalign{\vskip 2mm}  CLK & clock & O61735\\
\noalign{\vskip 2mm}  CRY & cryptochrome & O77059\\
\noalign{\vskip 2mm}  CWO & clockwork orange & Q9VGZ5\\
\noalign{\vskip 2mm}  CYC & cycle & O61734\\
\noalign{\vskip 2mm}  CYC/CLK & cycle-clock complex & \\
\noalign{\vskip 2mm}  DBT & double-time kinase & O76324\\
\noalign{\vskip 2mm}  DISC1 & death-inducing signalling complex 1 & \\
\noalign{\vskip 2mm}  DISC2 & death-inducing signalling complex 2 & \\
\noalign{\vskip 2mm}  EGF & epidermal growth factor & P01133\\
\noalign{\vskip 2mm}  EGFR & EGF receptor & P00533\\
\noalign{\vskip 2mm}  ERK & extracellular regulated kinase & P27361\\
\noalign{\vskip 2mm}  FKHR & forkhead box protein O1 & Q12778\\
\noalign{\vskip 2mm}  IKK & inhibitor of nuclear factor kappa-B kinase & O14920\\
\noalign{\vskip 2mm}  IL-1a & interleukin 1 $\alpha$ & P01583\\
\noalign{\vskip 2mm}  IL-1R & interleukin 1 receptor & P14778\\
\noalign{\vskip 2mm}  IL-1ra & interleukin 1 receptor antagonist & P18510\\
\noalign{\vskip 2mm}  IRS1 (S) & insulin receptor substrate 1 (Serine 636) & P35568\\
\noalign{\vskip 2mm}  IRS1 (Y) & insulin receptor substrate 1 (Tyrosine 896) & P35568\\
\noalign{\vskip 2mm}  JNK1 & c-Jun N-terminal kinase 1 & P45983\\
\noalign{\vskip 2mm}  MEK & MAPK ERK kinase & Q02750\\
\noalign{\vskip 2mm}  MEKK1 & MAPK/ERK kinase kinase 1 & Q13233\\
\noalign{\vskip 2mm}  MK2 & mitogen-activated protein kinase-activated protein kinase 2 & P49137\\
\noalign{\vskip 2mm}  MKK3/6 & dual specificity mitogen-activated protein kinase kinase 3/6 & P46734 / P52564\\
\noalign{\vskip 2mm}  MKK4/7 & dual specificity mitogen-activated protein kinase kinase 4/7 & P45985 / O14733\\
\noalign{\vskip 2mm}  NF-kB & nuclear factor kappa-B & P19838\\
\noalign{\vskip 2mm}  p38 & mitogen-activated protein kinase p38 & Q16539\\
\noalign{\vskip 2mm}  PDP1 & par-domain protein 1 & Q9TVY0\\
\noalign{\vskip 2mm}  PER & period & P07663\\
\noalign{\vskip 2mm}  PER/TIM-p & phosphorylated period-timeless complex & \\
\noalign{\vskip 2mm}  RAF & Raf & P04049\\
\noalign{\vskip 2mm}  RAS & Ras GTPase-activating protein & P01112\\
\noalign{\vskip 2mm}  TGF$\alpha$ & transforming growth factor $\alpha$ & P01135\\
\noalign{\vskip 2mm}  TNF$\alpha$ & tumour necrosis factor-$\alpha$ & P01375\\
\noalign{\vskip 2mm}  TNFR & TNF receptor & P19438\\
\noalign{\vskip 2mm}  TIM & timeless & P49021\\
\noalign{\vskip 2mm}  VRI & vrille & O18660\\[2mm]
\hline
%\end{tabular}
\end{longtable}
\end{center}

\clearpage
\section*{Normalizing experimental data for use with ANIMO}\label{suppl-sec:normalization}
Document S1 in the Supplemental Data of the work by~\cite{pathway-autocrine} contains
three tables, named {\sf Replicates}, {\sf Averages} and {\sf DPLSR dataset}.
The data we use to compare the results computed with ANIMO are based on the
values from the {\sf Averages} table.
In particular, we compute activity data by performing a normalization on a $0\dots 100$ scale using this formula
$$
v_{\mbox{\scriptsize norm}} = \frac{v}{v_{\mbox{\scriptsize max}}} \times 100
$$
where $v$ is the datum to be normalized taken from the column $v$ in the {\sf Averages} table,
$v_{\mbox{\scriptsize norm}}$ is the normalized value and
$v_{\mbox{\scriptsize max}}$ is the maximum value over the whole column $v$ in the {\sf Replicates} table.
For each series, we compute also the standard deviation using the triplicate measurements
present in table {\sf Replicates}. The standard deviation is also normalized using the formula presented for the average.
Each table produced with this process contains a subset of the columns from the {\sf Averages} table,
and refers to one treatment condition only. A column with time references is added to the table in first position.
Finally, a column named \emph{Number\_of\_levels} containing only the value $100$ (see the instructions
on page~\pageref{csv-import-format}) is added at the rightmost position.
The tables are all exported in .csv format to be used with ANIMO, and are included in the \emph{Model\_and\_data.zip}
file in the additional materials of the present work.

\clearpage
\section*{ANIMO and Timed Automata}\label{suppl-sec:animo-ta}

\subsection*{Timed Automata model}
The Timed Automata (\tas) model underlying an ANIMO network is generated whenever an analysis is requested by the user.
Starting from the network represented in the Cytoscape-based user interface, ANIMO automatically generates a \tas\ model
to be used with UPPAAL. The analysis result is then parsed and properly presented to the user, for example
as a graph of reactant activity levels. This workflow is described in Figure~\ref{fig:animo-sim-workflow}.

\begin{figure}[htb]
\begin{minipage}{\textwidth}
  \includegraphics[width=\textwidth]{images/animo_simulation_workflow}
\caption{The passages intercurring between the press of the \emph{Analyse network} button
and the generation of an activity level graph.
A simulation produces, for each selected node, a series of pairs {\sf ($t$, $a$)},
with $t$ a time instant along the simulation and $a$ an activity level. These data are
then parsed and translated into a graph.}\label{fig:animo-sim-workflow}
\end{minipage}
\end{figure}

Each \tas\ model generated by ANIMO contains one automaton for each interaction (activation or inhibition) in the network.
A \ta\ representing an interaction performs a cyclic series of steps, continuously updating
the target of the interaction it represents, and adapting the timing of the next update according to
the user-defined dynamics. Synchronizations between different automata occur when the activity level of a network component (e.g. ERK)
changes: this allows the automata depending on that component to update their time settings.

The abstract behaviour of the interaction $\mbox{MEK} \rightarrow \mbox{ERK}$ in the \tas\ model used in ANIMO is described in Figure~\ref{fig:ta-diagram}.
There, the activity levels of MEK and ERK are represented by variables called, respectively, $\mbox{MEK}_{\mbox{\scriptsize activity}}$
and $\mbox{ERK}_{\mbox{\scriptsize activity}}$. A more detailed description of the \tas\ model underlying ANIMO has been presented
at the IEEE conference BIBE 2012~\cite{animo-bibe}.

\tikzstyle{decision} = [rectangle, draw, fill=yellow!20, text width=.45\textwidth, text badly centered]
\tikzstyle{block} = [rectangle, draw, fill=blue!20,
    text width=.45\textwidth, text centered, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, thick, -latex']
\tikzstyle{cloud} = [draw, ellipse,fill=blue!20, minimum height=2em]
\def\svgwidth{.8\textwidth}
\definecolor{lowActivity}{rgb}{0.8,0,0}
\definecolor{highActivity}{rgb}{0,0.8,0}%{1,0.8,0}
\newsavebox{\mysquareLow}
\savebox{\mysquareLow}{%
  \raisebox{-0.08em}{%
    \textcolor{black}{%
      \rule{.7em}{.7em}%
    }%
    \hspace{-.65em}%
    \raisebox{.05em}{%
      \textcolor{lowActivity}{%
	\rule{.6em}{.6em}%
      }%
    }%
  }%
}
\newsavebox{\mysquareHigh}
\savebox{\mysquareHigh}{%
  \raisebox{-0.08em}{%
    \textcolor{black}{%
      \rule{.7em}{.7em}%
    }%
    \hspace{-.65em}%
    \raisebox{.05em}{%
      \textcolor{highActivity}{%
	\rule{.6em}{.6em}%
      }%
    }%
  }%
}

\begin{figure}[!ht]
\begin{minipage}{\textwidth}
\centering

\begin{tikzpicture}[node distance = 2cm, auto, scale=0.8, every node/.style={scale=0.8}]
    % Place nodes
    \node [block] (init) {
\begingroup
  \makeatletter
  \providecommand\color[2][]{%
    \errmessage{(Inkscape) Color is used for the text in Inkscape, but the package 'color.sty' is not loaded}
    \renewcommand\color[2][]{}%
  }
  \providecommand\transparent[1]{%
    \errmessage{(Inkscape) Transparency is used (non-zero) for the text in Inkscape, but the package 'transparent.sty'
is not loaded}
    \renewcommand\transparent[1]{}%
  }
  \providecommand\rotatebox[2]{#2}
  \ifx\svgwidth\undefined
    \setlength{\unitlength}{1974.66679688pt}
  \else
    \setlength{\unitlength}{\svgwidth}
  \fi
  \global\let\svgwidth\undefined
  \makeatother
  \begin{picture}(1,0.37)%26031697)%
    \put(0,0){\includegraphics[width=\unitlength]{images/mek_activity_time_tables2.pdf}}%
    \put(0.005,0.20182345){\color[rgb]{0,0,0}\makebox(0,0)[lb]{\smash{$R_{\mbox{\scriptsize lowerBound}}$}}}%
    \put(0.005,0.07442319){\color[rgb]{0,0,0}\makebox(0,0)[lb]{\smash{$R_{\mbox{\scriptsize upperBound}}$}}}%
    \put(0.45,0.325)%23815501)
	{\color[rgb]{0,0,0}\makebox(0,0)[lb]{\smash{$\mbox{MEK}_{\mbox{\scriptsize
activity}}$}}}%
  \end{picture}%
\endgroup
};
    \node [block, above of=init, node distance=2.5cm] (reset) {Reset the internal clock: ${\sf t} \!\!:= \!\!0$};
    \node [cloud, above of=reset, node distance=1.5cm] (start) {Start};
\node [block, below of=init, node distance=2.5cm] (choose) {$R$ will occur when
\raisebox{-0.3ex}{\pgfuseimage{lower-bound}} $\leq\!\!\! {\sf t}\!\!\! \leq$
\raisebox{-0.3ex}{\pgfuseimage{upper-bound}}
    };
    \node [block, below of=choose] (increase) {Increase $\mbox{ERK}_{\mbox{\scriptsize activity}}$
by $+1$ level};
    \node [block, below of=increase] (inform) {Inform interactions depending on $\mbox{ERK}_{\mbox{\scriptsize
activity}}$\\
    of the change};
    % Draw edges
    \path [line] (start) -- (reset);
    \path [line] (reset) -- (init);
    \path [line] (init) -- (choose);
    \path [line] (choose.west) -| +(-0.5, 2) node [pos=0.78, sloped, above, align=left] {$\mbox{MEK}_{\mbox{\scriptsize activity}}$ was changed} |- (init.west);
    \path [line] (choose) -- (increase);
    \path [line] (increase) -- (inform);
    \path [line] (inform.east) -| +(1, 4) |- (reset.east);
\end{tikzpicture}
\caption{Schematic overview of the steps taken during a simulation run by a Timed Automaton modelling an interaction $R$ that
increases $\mbox{ERK}_{\mbox{\scriptsize activity}}$ and depends only on $\mbox{MEK}_{\mbox{\scriptsize activity}}$.
In this example, MEK has 10 activity levels.\\
After resetting the internal clock {\sf t}, the automaton sets the time constraints for the interaction.
$\mbox{MEK}_{\mbox{\scriptsize activity}}$ is used as the index inside the time
tables $R_{\mbox{\scriptsize lowerBound}}$ and $R_{\mbox{\scriptsize upperBound}}$, which contain pre-computed lower- and upper-bounds
for the interaction timing.
Once the bounds have been identified, %the internal clock {\sf t} begins counting.
$R$ can occur when {\sf t} reaches a value
inside the continuous time interval~$[\,\usebox{\mysquareLow}, \usebox{\mysquareHigh}\,]$. When it occurs, $R$ increases the value of
$\mbox{ERK}_{\mbox{\scriptsize activity}}$ by $1$. All interactions that depend on
$\mbox{ERK}_{\mbox{\scriptsize activity}}$ are notified of the change (via a synchronization on a specific channel),
so that the associated time bounds are updated accordingly.
After resetting the clock {\sf t}, the process can restart.
If $\mbox{MEK}_{\mbox{\scriptsize activity}}$ was changed by another automaton before the occurrence of $R$,
the time bounds are updated according to the new activity level of MEK.}\label{fig:ta-diagram}
\end{minipage}
\end{figure}

\subsection*{Granularity of an ANIMO network node}
Figure~\ref{fig:levels} shows the differences between different choices for the
number of levels of a node. This allows to adapt a model to the quality of experimental data.

\begin{figure}[htpb]
\begin{minipage}{\textwidth}
\centering
\subfloat[\label{suppl:fig1-1level}]{\includegraphics[width=.45\textwidth]{images/JNK1_1level2}} \qquad
\subfloat[\label{suppl:fig1-10levels}]{\includegraphics[width=.45\textwidth]{images/JNK1_10levels2}} \\
\subfloat[\label{suppl:fig1-50levels}]{\includegraphics[width=.45\textwidth]{images/JNK1_50levels2}} \qquad
\subfloat[\label{suppl:fig1-100levels}]{\includegraphics[width=.45\textwidth]{images/JNK1_100levels2}}
\caption{Comparing different reactant granularity settings. {\bfseries (\protect\subref*{suppl:fig1-1level})} 2 levels, {\bfseries (\protect\subref*{suppl:fig1-10levels})} 10 levels,  {\bfseries (\protect\subref*{suppl:fig1-50levels})} 50 levels, {\bfseries (\protect\subref*{suppl:fig1-100levels})} 100 levels. The {\sf JNK1} series is computed from the model presented in Figure~\ref{fig:large-model-complete}, considering 100 ng/ml TNF$\alpha$ as treatment condition over a period of 60 minutes.}\label{fig:levels}
\end{minipage}
\end{figure}

\clearpage
\section*{Additional notes}

\subsection*{Simulating the day-night cycle}\label{suppl:repressilator}
The model presented in Figure~\ref{fig:drosophila-model} contains a node
labelled {\sf Day/Night}. That node abstracts our representation
of the cyclic alternation of day and night, which causes the variations
in cryptochrome ({\sf cry}): these oscillations allow the network
to synchronize to a time zone. Note that the network oscillates
also when the node {\sf cry} is not included in the model.

The alternation between day and night is represented in our model with a
repressilator-like~\cite{repressilator} subnetwork, as can be seen in Figure~\ref{fig:repressilator}.
In the model by~\cite{drosophila-ode-model} a specific function
was introduced in the equations to approximate the experimental data from~\cite{drosophila-cry-data}.

\def\reprGraph{\includegraphics[scale=0.15]{images/drosophila_model_repressilator}}
\newlength\reprGraphHeight
\setlength\reprGraphHeight{\heightof{\reprGraph}}
\begin{figure}[!htb]
\begin{minipage}{\textwidth}
  \centering
  \subfloat[\label{subfig:repressilator-model}]{\begin{minipage}[c][\reprGraphHeight]{0.35\textwidth}\begin{center}\reprGraph\end{center}\end{minipage}}\qquad
  \subfloat[\label{subfig:repressilator-graph}]{\begin{minipage}[c][\reprGraphHeight]{0.6\textwidth}\begin{center}\includegraphics[width=\textwidth]{images/CRY_oscillations}\end{center}\end{minipage}}
\caption{{\bf \protect\subref{subfig:repressilator-model}} The repressilator-like subnetwork used to represent the alternation
between day and night that cause the oscillations in {\sf CRY} concentrations in the
network modelled in Section~\ref{sec:animo-drosophila}.
{\bf \protect\subref{subfig:repressilator-graph}} A graph plotting the oscillations in {\sf CRY} along
a period of three days.}\label{fig:repressilator}
\end{minipage}
\end{figure}

\subsection*{ANIMO model of the Drosophila circadian clock}\label{suppl-sec:animo-drosophila}
We compared the simulation results from the ANIMO model presented in Figure~\ref{fig:drosophila-model}
with the ODE model described by~\cite{drosophila-ode-model}. The raw data coming from the
two models were aligned to have a roughly close initial point, and all amplitudes were normalized
following the procedure described in Suppl. Sect.~\ref{suppl-sec:normalization}. The results
of this comparison can be seen in Figure~\ref{suppl-fig:grafici-drosophila}.

\def\drosophilaGraphScale{0.0685}
\renewcommand*\thesubfigure{}
\begin{figure}[!htb]
\begin{minipage}{\textwidth}
  \centering
\begin{tabular}{ccc}
\subfloat[clk]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/clk}} &
\subfloat[CLK]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/CLK}} &
\subfloat[CRY]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/CRY}} \\
\subfloat[cwo]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/cwo}} &
\subfloat[CWO]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/CWO}} &
\subfloat[CYC/CLK]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/CYC-CLK}} \\
\subfloat[pdp1]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/pdp1}} &
\subfloat[PDP1]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/PDP1}} &
\subfloat[per]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/per}} \\
\subfloat[PER]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/PER}} &
\subfloat[PER/TIM]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/PER-TIM}} &
\subfloat[tim]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/tim}} \\
\subfloat[TIM]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/TIM}} &
\subfloat[vri]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/vri}} &
\subfloat[VRI]{\includegraphics[scale=\drosophilaGraphScale]{images/Circadian/VRI}}
\end{tabular}
\caption{Comparing the simulation results from the ANIMO and ODE models of
\emph{Drosophila Melanogaster} circadian clock. Blue lines ({\sf \_{}ANIMO} series)
represent data from the ANIMO model in Figure~\ref{fig:drosophila-model},
while red lines ({\sf \_{}ODE} series) represent data from the ODE model by~\cite{drosophila-ode-model}.\label{suppl-fig:grafici-drosophila}}
\end{minipage}
\end{figure}

Most of the molecules represented in the two models evolve with the same period and phase.
CLK and clk in the ANIMO model have
a small oscillation range (their values change by around 10\%\ during a simulation),
so their behaviour match the continuous model less precisely.
% The parameters of the ANIMO model used to represent the \emph{Drosophila Melanogaster} circadian clock in Section~\ref{sec:animo-drosophila}
% are given in Tables~\ref{tab:drosophila-model-reactants} and~\ref{tab:drosophila-model-reactions}.
%
%
%
% \begin{table}[htbp]
% \begin{minipage}{\textwidth}
% \begin{center}
% \processtable{The settings for the nodes in the model from Section~\ref{sec:animo-drosophila}.\label{tab:drosophila-model-reactants}}
% % {\begin{tabular}{llllll}%|c|c|c|c|c|c|}
% % \ \\
% % \hline\noalign{\vskip 2mm}
% %   \multirow{2}{*}{{\bfseries Name}} & {\bfseries Total act.} & {\bfseries Initial act.} & \multirow{2}{*}{{\bfseries Molecule type}} &
% % \multirow{2}{*}{{\bfseries Enabled?}} & \multirow{2}{*}{{\bfseries Plotted?}}\\
% % & {\bfseries levels} & {\bfseries level} & & & \\[2mm]
% % \hline\noalign{\vskip 2mm}
% % clk & 100 & 98 & Other & Yes & Yes \\[5mm]
% % CLK & 100 & 100 & Other & Yes & Yes \\[5mm]
% % cry & 100 & 100 & Other & Yes & Yes \\[5mm]
% % CRY & 100 & 80 & Other & Yes & Yes \\[5mm]
% % cwo & 100 & 0 & Other & Yes & Yes \\[5mm]
% % CWO & 100 & 17 & Other & Yes & Yes \\[5mm]
% % CYC & 100 & 100 & Other & Yes & No \\[5mm]
% % CYC/CLK & 100 & 0 & Other & Yes & Yes \\[5mm]
% % DBT & 100 & 100 & Other & Yes & No \\[5mm]
% % pdp1 & 100 & 0 & Other & Yes & Yes \\[5mm]
% % PDP1 & 100 & 49 & Other & Yes & Yes \\[5mm]
% % PER & 100 & 5 & Other & Yes & Yes \\[5mm]
% % per & 100 & 0 & Other & Yes & Yes \\[5mm]
% % PER/TIM-p & 100 & 46 & Other & Yes & Yes \\[5mm]
% % tim & 100 & 0 & Other & Yes & Yes \\[5mm]
% % TIM & 100 & 5 & Other & Yes & Yes \\[5mm]
% % VRI & 100 & 39 & Other & Yes & Yes \\[5mm]
% % vri & 100 & 0 & Other & Yes & Yes \\[5mm]
% % (1) & 100 & 100 & Other & Yes & No \\[5mm]
% % (2) & 100 & 69 & Other & Yes & No \\[5mm]
% % (3) & 100 & 100 & Other & Yes & No \\[5mm]
% % (4) & 100 & 0 & Other & Yes & No \\[5mm]
% % (5) & 100 & 0 & Other & Yes & No \\[5mm]
% % (6) & 100 & 100 & Other & Yes & No \\[2mm]
% % \hline
% % \end{tabular}
% {\begin{tabular}{lll||lll}
% \ \\
% \hline%\noalign{\vskip 2mm}
% & & & & & \\[-1mm]
%   \multirow{2}{*}{{\bfseries Name}} & {\bfseries Total act.} & {\bfseries Initial act.} & \multirow{2}{*}{{\bfseries Name}} &
% {\bfseries Total act.} & {\bfseries Initial act.}\\
% & {\bfseries levels} & {\bfseries level} & & {\bfseries levels} & {\bfseries level} \\[2mm]
% \hline%\noalign{\vskip 5mm}
% & & & & & \\
% clk & 100 & 98 & per & 100 & 0 \\[5mm]
% CLK & 100 & 100 & PER/TIM-p & 100 & 46 \\[5mm]
% cry & 100 & 100 & tim & 100 & 0 \\[5mm]
% CRY & 100 & 80 & TIM & 100 & 5 \\[5mm]
% cwo & 100 & 0 & VRI & 100 & 39 \\[5mm]
% CWO & 100 & 17 & vri & 100 & 0 \\[5mm]
% CYC & 100 & 100 & (1) & 100 & 100 \\[5mm]
% CYC/CLK & 100 & 0 & (2) & 100 & 69 \\[5mm]
% DBT & 100 & 100 & (3) & 100 & 100 \\[5mm]
% pdp1 & 100 & 0 & (4) & 100 & 0 \\[5mm]
% PDP1 & 100 & 49 & (5) & 100 & 0 \\[5mm]
% PER & 100 & 5 & (6) & 100 & 100 \\[2mm]
% \hline
% \end{tabular}
% }{}
% \end{center}
% \end{minipage}
% \end{table}
%
%
% \begin{table}[!ht]
% \begin{minipage}{\textwidth}
% \processtable{The settings for the edges (interactions) in the model from Section~\ref{sec:animo-drosophila}.
% $\rightarrow$ indicates activation, while $\dashv$ stands for inhibition.\label{tab:drosophila-model-reactions}}
% {\begin{tabular}{lll||lll}
% \ \\
% \hline%\noalign{\vskip 2mm}
% & & & & & \\[-1mm]
%   {\bfseries Interaction} & {\bfseries Scenario} & {\bfseries Param. value} & {\bfseries Interaction} & {\bfseries Scenario} & {\bfseries Param. value}\\[2mm]
% \hline
% & & & \\
% % \noalign{\vskip 2mm}  DBT $\rightarrow$ PER/TIM-p & 1 & 0.001 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  PER/TIM-p $\dashv$ CYC/CLK & 1 & 0.32 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  CYC $\rightarrow$ CYC/CLK & 1 & 0.02 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CLK $\rightarrow$ CYC/CLK & 1 & 0.072 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  clk $\rightarrow$ CLK & 1 & 0.024 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CYC/CLK $\rightarrow$ per & 1 & 0.031 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CYC/CLK $\rightarrow$ tim & 1 & 0.015 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CYC/CLK $\rightarrow$ cwo & 1 & 0.1 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CYC/CLK $\rightarrow$ pdp1 & 1 & 0.0375 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CYC/CLK $\rightarrow$ vri & 1 & 0.03 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  TIM $\rightarrow$ PER/TIM-p & 1 & 0.069 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  PER $\rightarrow$ PER/TIM-p & 1 & 0.053 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  per $\rightarrow$ PER & 1 & 0.0075 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  cwo $\rightarrow$ CWO & 1 & 0.03 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CWO $\dashv$ cwo & 1 & 0.112 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  vri $\rightarrow$ VRI & 1 & 0.026 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  VRI $\dashv$ clk & 1 & 0.0098 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  CWO $\dashv$ per & 1 & 0.05 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  CWO $\dashv$ vri & 1 & 0.024 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  CWO $\dashv$ tim & 1 & 0.02 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  CWO $\dashv$ pdp1 & 1 & 0.0384 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  tim $\rightarrow$ TIM & 1 & 0.056 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  pdp1 $\rightarrow$ PDP1 & 1 & 0.042 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  PDP1 $\rightarrow$ clk & 1 & 0.01 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  cry $\rightarrow$ CRY & 2 & 0.9733 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  CRY $\dashv$ TIM & 1 & 0.048 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  cry $\dashv$ (2) & 1 & 0.10512 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  (2) $\rightarrow$ (3) & 1 & 0.10512 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  (3) $\dashv$ (4) & 1 & 0.10512 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  (4) $\rightarrow$ (5) & 1 & 0.10512 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  (5) $\dashv$ (6) & 1 & 0.10512 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  (6) $\rightarrow$ cry & 1 & 0.10512 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  (1) $\dashv$ cry & 1 & 0.0257 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  (1) $\rightarrow$ (2) & 1 & 0.0.0518 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  (1) $\dashv$ (3) & 1 & 0.0257 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  (1) $\rightarrow$ (4) & 1 & 0.0518 & Activation\\[5mm]
% % \noalign{\vskip 2mm}  (1) $\dashv$ (5) & 1 & 0.0257 & Inhibition\\[5mm]
% % \noalign{\vskip 2mm}  (1) $\rightarrow$ (6) & 1 & 0.0518 & Activation\\[2mm]
% DBT $\rightarrow$ PER/TIM-p & 1 & 0.001 & CWO $\dashv$ tim & 1 & 0.02 \\[5mm]
% PER/TIM-p $\dashv$ CYC/CLK & 1 & 0.32 & CWO $\dashv$ pdp1 & 1 & 0.0384 \\[5mm]
% CYC $\rightarrow$ CYC/CLK & 1 & 0.02 & tim $\rightarrow$ TIM & 1 & 0.056 \\[5mm]
% CLK $\rightarrow$ CYC/CLK & 1 & 0.072 & pdp1 $\rightarrow$ PDP1 & 1 & 0.042 \\[5mm]
% clk $\rightarrow$ CLK & 1 & 0.024 & PDP1 $\rightarrow$ clk & 1 & 0.01 \\[5mm]
% CYC/CLK $\rightarrow$ per & 1 & 0.031 & cry $\rightarrow$ CRY & 2 & 0.9733 \\[5mm]
% CYC/CLK $\rightarrow$ tim & 1 & 0.015 & CRY $\dashv$ TIM & 1 & 0.048 \\[5mm]
% CYC/CLK $\rightarrow$ cwo & 1 & 0.1 & cry $\dashv$ (2) & 1 & 0.10512 \\[5mm]
% CYC/CLK $\rightarrow$ pdp1 & 1 & 0.0375 & (2) $\rightarrow$ (3) & 1 & 0.10512 \\[5mm]
% CYC/CLK $\rightarrow$ vri & 1 & 0.03 & (3) $\dashv$ (4) & 1 & 0.10512 \\[5mm]
% TIM $\rightarrow$ PER/TIM-p & 1 & 0.069 & (4) $\rightarrow$ (5) & 1 & 0.10512 \\[5mm]
% PER $\rightarrow$ PER/TIM-p & 1 & 0.053 & (5) $\dashv$ (6) & 1 & 0.10512 \\[5mm]
% per $\rightarrow$ PER & 1 & 0.0075 & (6) $\rightarrow$ cry & 1 & 0.10512 \\[5mm]
% cwo $\rightarrow$ CWO & 1 & 0.03 & (1) $\dashv$ cry & 1 & 0.0257 \\[5mm]
% CWO $\dashv$ cwo & 1 & 0.112 & (1) $\rightarrow$ (2) & 1 & 0.0.0518 \\[5mm]
% vri $\rightarrow$ VRI & 1 & 0.026 & (1) $\dashv$ (3) & 1 & 0.0257 \\[5mm]
% VRI $\dashv$ clk & 1 & 0.0098 & (1) $\rightarrow$ (4) & 1 & 0.0518 \\[5mm]
% CWO $\dashv$ per & 1 & 0.05 & (1) $\dashv$ (5) & 1 & 0.0257 \\[5mm]
% CWO $\dashv$ vri & 1 & 0.024 & (1) $\rightarrow$ (6) & 1 & 0.0518 \\[2mm]
% \hline
% \end{tabular}}{}
% \end{minipage}
% \end{table}\vspace{-2ex}

\subsection*{Note on the parameters in the TNF$\alpha$-EGF model}\label{suppl:parameters-tnf-egf}
The parameters in the model in Figure~\ref{fig:large-model-complete}
have been set by fitting the model to the experimental data for conditions with 100 ng/ml TNF$\alpha$.
In the model we have set the starting level of TNF$\alpha$ at 100 out of 100 for these conditions.
This level is a dimensionless quantity that indicates the maximum activity level in the data set.
We found that setting the initial level of TNF$\alpha$ at level 8 out of 100 gave slightly better results for the
condition with 5 ng/ml TNF$\alpha$ than level 5 out of 100. We believe that this has to do with the fact that
100 ng/ml is a highly supra-physiological concentration of TNF$\alpha$, that will rapidly cause activation of all
receptors present. Fitting the model to this experimental condition may have resulted in slight deviations
in the parameter values. Nevertheless, the modelling results illustrate that building a model with basic
kinetic rate laws can give useful predictions over a range of concentrations. Figures~\ref{fig:large-model-graph3} and~\ref{fig:large-model-graph4}
show the modelling results with TNF set at 8 out of 100.

\clearpage
\section*{Comparison between ANIMO and other modelling tools}\label{suppl:comparison-table}
Different formalisms are in use in the field of computational
modelling of biological systems, each with their specific characteristics.
Many of these formalisms have been implemented into
software tools to support modelling efforts. In order to compare
ANIMO with existing tools, we have selected a number of mathematical formalisms,
each connected to a supporting tool. With an emphasis on the modelling
process rather than the final model, we compared these tools on
the basis of the following parameters:

\begin{enumerate}
  \item {\bf Hidden formalism:} a knowledge of the underlying formalism is not required in order to use the tool
  \item {\bf Visual modelling:} the tool allows the user to model using a visual interface, and is not exclusively
      founded on formula-, text- or table-based input forms
  \item {\bf Qualitative parameters:} parameters for reactions can be input as approximated estimations, and not exclusively as numbers
  \item {\bf Tight coupling with topology:} models are tightly and clearly coupled to the networks they represent, showing the visual
      representation of the model in a shape similar or comparable to the representation currently used by biologists
      for signalling pathways
  \item {\bf User-chosen granularity:} if discretization is applied during the modelling process, the user can change the granularity
      with which such discretization is made, possibly for each component of the model separately
\end{enumerate}
Table~\ref{tab:tool-comparison} shows the comparison between ANIMO and the selected tools.

\newcolumntype{x}[1]{%
>{\centering\hspace{0pt}}p{#1}}%
\savenotes
\begin{table}[!hbt]
\begin{minipage}{\textwidth}
\processtable{Comparison between ANIMO and some existing approaches to modelling biological systems.
A ``Yes'' under a column indicates that the modelling tool (mostly) fulfils the parameter, ``No'' indicates very limited or no fulfilment.\\
\label{tab:tool-comparison}}
{\begin{tabular}{p{4cm}p{1.8cm}p{1.8cm}p{1.5cm}p{1.8cm}p{1.8cm}p{2cm}}
\toprule
{\bfseries Tool} & {\bfseries Formalism} & {\bfseries Hidden formalism} & {\bfseries Visual modelling} & {\bfseries Qualitative parameters} & {\bfseries Tight coupling with topology} & {\bfseries User-chosen granularity}  \\[5mm]
\midrule
\noalign{\vskip 2mm} ANIMO~\cite{animo-bibe} & Timed \ \ \ Automata & Yes & Yes & Yes & Yes & Yes
      \\[5mm]
\noalign{\vskip 2mm} Bio-PEPA Workbench~\cite{biopepa-interface} & Bio-PEPA & No & No & No & No & Yes
      \\[5mm]
\noalign{\vskip 2mm} Cell Illustrator~\cite{cell-illustrator} & Petri Nets & Yes & Yes & No & Yes & No
     \\[5mm]
\noalign{\vskip 2mm} COPASI~\cite{copasi} & ODE, stochastic models & No & No & No & No & No
      \\[5mm]
\noalign{\vskip 2mm} COSBI LAB~$^1$
 & BlenX & Yes & Yes & No & Yes & No
      \\[5mm]
\noalign{\vskip 2mm} GINsim~\cite{ginsim} & Boolean Networks & No & Yes & Yes & Yes & Yes~$^2$
      \\[5mm]
\noalign{\vskip 2mm} GNA~\cite{gna}    & ODE & No & Yes & Yes & Yes & No~$^3$
      \\
\noalign{\vskip 2mm} Rhapsody~$^4$
 & Statecharts & No & Yes & Yes & No~$^5$ & No
      \\[5mm]
\botrule
\end{tabular}}{$^1$ {COSBILab} web page \url{http://www.cosbi.eu/index.php/research/cosbi-lab}\\
$^2$ The user can choose the number of levels for each reactant, allowing to define
multi-level models based on Boolean reaction dynamics.\\
$^3$ When discretizing an ODE model, the granularity depends on the mathematical
features of the model, and not directly on the user's choice.\\
$^4$ {IBM Rational Rhapsody} web page \url{http://www-01.ibm.com/software/rational/products/rhapsody/designer}\\
$^5$ Statecharts represent more closely the so-called
\emph{transition system} of the model as opposed to the components and interactions occurring among them.}
\end{minipage}
\end{table}

\clearpage
\addcontentsline{toc}{section}{References}
\begin{thebibliography}{}

\bibitem[Chaouiya et~al., 2003]{ginsim}
Chaouiya, C., Remy, E., Moss\'{e}, B., and Thieffry, D. (2003).
\newblock Qualitative analysis of regulatory graphs: A computational tool based
  on a discrete formal framework.
\newblock In Benvenuti, L., De~Santis, A., and Farina, L., editors, {\em
  Positive Systems}, volume 294 of {\em Lecture Notes in Control and
  Information Sciences}, pages 830--832. Springer Berlin / Heidelberg.

\bibitem[Ciocchetta et~al., 2009]{biopepa-interface}
Ciocchetta, F., Duguid, A., Gilmore, S., Guerriero, M.~L., and Hillston, J.
  (2009).
\newblock {The Bio-PEPA Tool Suite}.
\newblock {\em International Conference on Quantitative Evaluation of Systems},
  pages 309--310.

\bibitem[de~Jong et~al., 2003]{gna}
de~Jong, H., Geiselmann, J., Hernandez, C., and Page, M. (2003).
\newblock {Genetic Network Analyzer: qualitative simulation of genetic
  regulatory networks}.
\newblock {\em Bioinformatics}, 19(3):336--344.

\bibitem[Elowitz, 2000]{repressilator}
Elowitz, Michael~B.Leibler, S. (2000).
\newblock A synthetic oscillatory network of transcriptional regulators.
\newblock {\em Nature}, 403(6767):335.

\bibitem[Fathallah-Shaykh et~al., 2009]{drosophila-ode-model}
Fathallah-Shaykh, H.~M., Bona, J.~L., and Kadener, S. (2009).
\newblock Mathematical model of the drosophila circadian clock: Loop regulation
  and transcriptional integration.
\newblock {\em Biophysical Journal}, 97(9):2399 -- 2408.

\bibitem[Janes et~al., 2006]{pathway-autocrine}
Janes, K.~A., Gaudet, S., Albeck, J.~G., Nielsen, U.~B., Lauffenburger, D.~A.,
  and Sorger, P.~K. (2006).
\newblock The response of human epithelial cells to {TNF} involves an inducible
  autocrine cascade.
\newblock {\em Cell}, 124(6):1225--1239.

\bibitem[Kadener et~al., 2007]{drosophila-cry-data}
Kadener, S., Stoleru, D., McDonald, M., Nawathean, P., and Rosbash, M. (2007).
\newblock Clockwork orange is a transcriptional repressor and a new drosophila
  circadian pacemaker component.
\newblock {\em Genes \& Development}, 21(13):1675--1686.

\bibitem[Killcoyne et~al., 2009]{cytoscape}
Killcoyne, S., Carter, G.~W., Smith, J., and Boyle, J. (2009).
\newblock {Cytoscape: a community-based framework for network modeling.}
\newblock {\em Methods in molecular biology (Clifton, N.J.)}, 563:219--239.

\bibitem[Larsen et~al., 1997]{uppaal}
Larsen, K.~G., Pettersson, P., and Yi, W. (1997).
\newblock {UPPAAL} in a nutshell.
\newblock {\em International Journal on Software Tools for Technology Transfer
  (STTT)}, 1:134--152.

\bibitem[Mendes et~al., 2009]{copasi}
Mendes, P., Hoops, S., Sahle, S., Gauges, R., Dada, J., and Kummer, U. (2009).
\newblock {Computational modeling of biochemical networks using COPASI systems
  biology}.
\newblock volume 500 of {\em Methods in Molecular Biology}, chapter~2, pages
  17--59. Humana Press, Totowa, NJ.

\bibitem[Nagasaki et~al., 2011]{cell-illustrator}
Nagasaki, M., Saito, A., Jeong, E., Li, C., Kojima, K., Ikeda, E., and Miyano,
  S. (2011).
\newblock Cell illustrator 4.0: a computational platform for systems biology.
\newblock {\em Stud Health Technol Inform}, 162:160--81.

\bibitem[Schivo et~al., 2012]{animo-bibe}
Schivo, S., Scholma, J., Wanders, B., {Urquidi Camacho}, R.~A., van~der Vet,
  P.~E., Karperien, M., Langerak, R., van~de Pol, J., and Post, J.~N. (2012).
\newblock Modelling biological pathway dynamics with {Timed Automata}.
\newblock In {\em 2012 IEEE 12th International Conference on Bioinformatics \&
  Bioengineering (BIBE)}, pages 447--453, Los Alamitos, CA, USA. IEEE Computer
  Society.

\end{thebibliography}

\clearpage

\section*{Supplementary figures}\label{sec:supplementary-figures}
\addcontentsline{toc}{section}{Supplementary figures}

\begin{figure}[htpb]
\begin{minipage}{\textwidth}
\centering
  \includegraphics[width=.7\textwidth]{images/large_network_tnfa2}
\caption{The model for the TNF$\alpha$ pathway in isolation. Node colours represent the activity level of the
corresponding modelled reactants at time $t = 10$ minutes after a stimulation of 100 ng/ml TNF$\alpha$.}\label{fig:large-model-tnf}
\end{minipage}
\end{figure}

\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\centering
  \includegraphics[width=.7\textwidth]{images/large_network_egf3}
\caption{The model for the EGF pathway in isolation. Node colours represent the
activity level of the corresponding modelled reactants at time $t = 5$ minutes after
a stimulation of 100 ng/ml EGF.}\label{fig:large-model-egf}
\end{minipage}
\end{figure}

\def\largeModelScale{0.18}%0.15}%0.155}%0.27}
\def\legendScalaColori{0.21}%0.18}%0.21}
\def\legendScalaForme{0.21}%0.18}%0.21}
\def\scalaGrafici{0.0709}%0.13}
\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\centering
  \subfloat{\includegraphics[width=\textwidth]{images/large_network_hypotheses7}}\\
\subfloat{\includegraphics[scale=\legendScalaColori]{images/legenda_forme}} \subfloat{\includegraphics[scale=\legendScalaForme]{images/legenda_colori}}
\caption{The merged model for the TNF$\alpha$-EGF pathway in which
the two hypotheses are highlighted. The first hypothesis is the dependence IL-1$\alpha$ expression on the
combined activity of ERK and JNK1. The second hypothesis assumes an as yet unidentified protein (Hyp2) to link EGFR to MEKK1.
Node colours represent initial activity levels.}\label{fig:large-model-hypotheses}
\end{minipage}
\end{figure}

\renewcommand*\thesubfigure{}
\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\begin{tabularx}{\textwidth}{XXX}
\subfloat[MK2 (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/MK2}} &
\subfloat[JNK1 (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/JNK1}} &
\subfloat[IKK (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/IKK_120}} \\
\subfloat[IKK (24 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/IKK_1440}} &
\subfloat[caspase-8 (24 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/casp8}} &
\subfloat[caspase-3 (24 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/casp3}} \\
\subfloat[MEK (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/MEK}} &
\subfloat[ERK (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/ERK}} &
\subfloat[Akt (24 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/Akt}} \\
\subfloat[IRS-1(S) (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/IRS1s_120}} &
\subfloat[IRS-1(Y) (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/IRS1y_120}} &
\subfloat[FKHR (12 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100/FKHR_720}}
\end{tabularx}
\caption{Comparison between the ANIMO model in Figure~\ref{fig:large-model-complete} and experimental data. Treatment condition: 100 ng/ml TNF$\alpha$.
In order to ease the comparison for earlier responses, the time span for those cases is less than 24 hours.}\label{fig:differences1}
\end{minipage}
\end{figure}

\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\begin{tabularx}{\textwidth}{XXX}
\subfloat[MK2 (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/MK2}} &
\subfloat[JNK1 (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/JNK1}} &
\subfloat[IKK (24 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/IKK}} \\
\subfloat[caspase-8 (24 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/casp8}} &
\subfloat[caspase-3 (24 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/casp3}} &
\subfloat[MEK (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/MEK}} \\
\subfloat[ERK (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/ERK}} &
\subfloat[Akt (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/Akt_120}} &
\subfloat[Akt (24 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/Akt_1440}} \\
\subfloat[IRS-1(S) (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/IRS1s_120}} &
\subfloat[IRS-1(Y) (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/IRS1y_120}} &
\subfloat[IRS-1(Y) (12 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/IRS1y_720}} \\
\subfloat[FKHR (2 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/FKHR_120}} &
\subfloat[FKHR (12 hours)]{\includegraphics[width=0.29\textwidth]{images/EGF100/FKHR_720}}
\end{tabularx}
\caption{Comparison between the ANIMO model in Figure~\ref{fig:large-model-complete} and experimental data. Treatment condition: 100 ng/ml EGF.
In order to ease the comparison for earlier responses, the time span for those cases is less than 24 hours.}\label{fig:differences2}
\end{minipage}
\end{figure}

\begin{figure}[!tpb]
\begin{minipage}{\textwidth}
\begin{tabularx}{\textwidth}{XXX}
\subfloat[MK2 (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/MK2}} &
\subfloat[JNK1 (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/JNK1}} &
\subfloat[IKK (24 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/IKK}} \\
\subfloat[caspase-8 (24 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/casp8}} &
\subfloat[caspase-3 (24 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/casp3}} &
\subfloat[MEK (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/MEK}} \\
\subfloat[ERK (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/ERK}} &
\subfloat[Akt (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/Akt}} &
\subfloat[IRS-1(S) (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/IRS1s_120}} \\
\subfloat[IRS-1(S) (12 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/IRS1s_720}} &
\subfloat[IRS-1(Y) (2 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/IRS1y}} &
\subfloat[FKHR (4 hours)]{\includegraphics[width=0.29\textwidth]{images/TNFa100_EGF100/FKHR_240}}
\end{tabularx}
\caption{Comparison between the ANIMO model in Figure~\ref{fig:large-model-complete} and experimental data. Treatment condition: 100 ng/ml TNF$\alpha$ + 100 ng/ml EGF.
In order to ease the comparison for earlier responses, the time span for those cases is less than 24 hours.}\label{fig:differences3}
\end{minipage}
\end{figure}

\end{document}